### 2024-11-12 doc/paper.qmd
Initial commit

---
title: |
  | Earnings Management and Investor Protection:
  | Accounting Reading Group - Assignment III\vspace{1cm}
author:
  - name: Melisa Mazaeva
    email: melisa.mazaeva@student.hu-berlin.de
    affiliations:
      - Humboldt-Universität zu Berlin  
date: today
date-format: MMM D, YYYY [\vspace{1cm}]
abstract: |
  | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide an infrastructure for open science-oriented empirical projects. Leveraging Worldscope data on financial information, this repository showcases a reproducible workflow that integrates Python scripts for data analysis. The project’s output demonstrates a comprehensive application of skills to replicate and extend the findings from the seminal paper by Leuz, Nanda, and Wysocki (2003), particularly in providing descriptive statistics for four individual earnings management measures, as well as an aggregate earnings management score across various countries. In doing so, it documents and discusses the research design choices made and the variations between the original and reproduced results. This code base, adapted from TREAT, provides an overview of how the template can be used for this specific project and how to structure a reproducible empirical research project.
  | \vspace{6cm}
bibliography: references.bib
biblio-style: apsr
format:
  pdf:
    documentclass: article
    number-sections: true
    toc: true
fig_caption: yes
fontsize: 11pt
ident: yes
always_allow_html: yes
number-sections: true 
header-includes:
  - \usepackage[nolists]{endfloat}    
  - \usepackage{setspace}\doublespacing
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
  - \usepackage[hang,flushmargin]{footmisc}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=24pt,font=bf}
  - \usepackage{array}
  - \usepackage{threeparttable}
  - \usepackage{adjustbox}
  - \usepackage{graphicx}
  - \usepackage{csquotes}
  - \usepackage{indentfirst}  # Added this line to ensure the first paragraph is indented for better readability
  - \usepackage[margin=1in]{geometry}
---

\pagebreak

```{python}
#| echo: false
#| output: false

import pickle
import pandas as pd

# Load the results from the existing pickle file for table_1
with open('../data/generated/table_1.pickle', 'rb') as f:
    results = pickle.load(f)

# Load the results from the new pickle file for final_table and summary_stats
with open('../output/em_results.pickle', 'rb') as f:
    em_results = pickle.load(f)

def escape_special_characters(df):
    df.columns = [col.replace('_', '\\_').replace('#', '\\#') for col in df.columns]
    return df

# Custom formatter function
def custom_float_format(x):
    return f"{x:.3f}".lstrip('0') if pd.notnull(x) else ""

def prep_latex_table(df, caption=None, label=None, fontsize='\\fontsize{11}{13}\\selectfont', description=None):
    df = escape_special_characters(df.reset_index(drop=True))  # Drop the index column from output tables for better readability
    num_columns = len(df.columns)
    column_format = 'l' + 'r' * (num_columns - 1)

    latex_table = df.to_latex(
        column_format=column_format,
        index=False, 
        float_format=custom_float_format,
        escape=False  # Important to disable escaping so LaTeX commands in column names are respected
    )

    # Construct description to the table
    full_caption = f"{caption}\\\\\n\\textnormal{{{description}}}" if description else caption

    latex_table_lines = [
        "\\begin{table}[htbp]",
        "\\centering",
        fontsize,  # Set font size for the table
        latex_table,
        f"\\caption{{{full_caption}}}" if full_caption else "",
        f"\\label{{{label}}}" if label else "",
        "\\end{table}"
    ]
    return "\n".join(line for line in latex_table_lines if line)
```

\pagebreak

# List of Abbreviations
\setlength{\parindent}{0em}

**CFO**: Operating Cash Flow  
**EM**: Earnings Management  
**IDE**: Integrated Development Environment  
**SIC**: Standard Industrial Classification  
**UK**: United Kingdom  
**US**: United States  
**WRDS**: Wharton Research Data Services

\setlength{\parindent}{4em}

\pagebreak

# Introduction {#sec-introduction}
The aim of this paper is to illustrate the use of open science tools in empirical accounting research.

The original study by @Leuz_2003 examines systematic differences in earnings management (EM) across 31 countries and proposes that these differences are influenced by the level of investor protection. Specifically, the paper argues that stronger investor protection reduces earnings management by limiting insiders’ ability to acquire private control benefits.

This paper presents the replication of Table 2 Panel A from the seminal study by @Leuz_2003. The original table provides descriptive statistics for four individual earnings management measures and the aggregate earnings management score across various countries, highlighting significant differences in EM practices between Continental Europe, Asia, and Anglo-American countries.

The project uses data from the Worldscope Database via Wharton Research Data Services (WRDS), focusing on fiscal years 1990 to 1999 and covering multiple countries. The replication aims to follow the research design by @Leuz_2003 as closely as possible with the available data. It employs four individual earnings management measures: earnings smoothing using accruals (EM1), the correlation between changes in accruals and cash flows (EM2), the magnitude of accruals (EM3), and the ratio of small profits to small losses (EM4). These measures are combined to create an aggregate earnings management score, providing a comprehensive overview of EM practices.

The analysis includes pulling and filtering the relevant financial data, identifying relevant companies and countries based on the authors’ requirements, calculating the EM measures for each country, and examining the relationship between each country’s aggregate earnings management score and its ranking. Explicit assumptions were made whenever the paper by @Leuz_2003 was unclear on how to proceed. The entire Python computation code for EM measures is available in `code/python/do_analysis.py` for detailed review if necessary. This paper focuses on the replication process and discussion of the results.

The replicated table is shown below, followed by the research design choices and assumptions in @sec-research_design_assumptions, documentation of the replication steps in @sec-replication_steps, a detailed results comparison in @sec-results, and concluding remarks in @sec-conclusion.

Table 1 provides descriptive statistics for the four individual earnings management measures as well as the aggregate earnings management score, where countries are sorted in descending order based on their aggregate score. The discussion of replication findings is presented in @sec-results.
```{python}
#| label: final_combined_table
#| echo: false
#| output: asis

# Rename the columns to ensure consistency with paper
em_results['final_combined_table'].rename(columns={
    'item6026': 'Country',
    'EM1': 'EM1 (−)',
    'EM2': 'EM2 (−)',
    'EM3': 'EM3 (+)',
    'EM4': 'EM4 (+)',
    'Aggregate_EM_Score': 'Aggregate EM score'
}, inplace=True)

print(prep_latex_table(
    em_results['final_combined_table'], 
    caption='Replicated table - Country scores for earnings management measures (Sorted by aggregate earnings management)',
    label='tab:final_combined_table',
    description = (
    "\\fontsize{10pt}{12pt}\\selectfont " 
    "The variables are computed from 123,469 firm-year observations for fiscal years 1990 to 1999 across 31 countries and 18,040 non-financial firms. Data are obtained from the Worldscope Database (July 2024 version). The definitions and calculations of EM1-EM4, as well as the aggregate score, are thoroughly presented in Section 4. The sign in the column heading indicates whether higher scores for the respective EM measure imply more earnings management (+) or less earnings management (−)."
    )

))
```

# Research Design Choices and Assumptions {#sec-research_design_assumptions}
The aim of Assignment III is to replicate a specific empirical table (Table 2 Panel A) from the seminal paper by @Leuz_2003. This table involves calculating the EM measures for firms across various countries over a defined period and examining the relationship between these measures and investor protection. The replication process includes data loading, preparation, cleaning, and normalization, followed by the application of statistical methods to compute and interpret financial metrics. For Assignment III, I pulled data from the Worldscope database through WRDS and used the Python programming language to carry out the empirical analysis. Visual Studio Code was used as the Integrated Development Environment (IDE) for writing, debugging, and optimizing the Python code.

The replication is based on data pulled from the Worldscope database, specifically from the `wrds_ws_company` and `wrds_ws_funda` tables, which were merged for the analysis. The first table provides company profile information, including items such as ISIN, Worldscope Identifiers, company name, and the country where the company is domiciled [@WRDS_WS_Company_2024]. The latter table contains fundamental annual data at the company-year level, including items such as total assets, net income, and other relevant financial variables [@WRDS_WS_Funda_2024].

Following @Leuz_2003, I focus the analysis on companies across various countries, ensuring that the data accurately reflects the fiscal years 1990 to 1999 as specified in the original study. The replication aims to mirror the research design as closely as possible with the available data.

In addition, I impose the following assumptions to ensure clarity and consistency where the paper by @Leuz_2003 does not provide explicit guidance:

1.	The original study references the November 2000 version of the Worldscope Database. However, the data used for this analysis represents the latest available version, updated in July 2024, with quarterly frequency updates [@WRDS_2024]. Due to potential adjustments and updates made to the database since November 2000, there may be differences between the databases that could affect the results. For example, companies may restate financials after the original reporting period, so these restatements are reflected in the later database version rather than the historical one. Moreover, the data vendor Refinitiv regularly updates its databases to correct errors and add new information, which may be included in the later data but not in the November 2000 snapshot.
2.	The original paper outlines key terms that will be used in this project to ensure consistency and accuracy in the replication. @Leuz_2003 define earnings management as the manipulation of a firm’s reported economic performance by insiders to deceive certain stakeholders or to affect contractual outcomes. The authors describe investor protection as a key institutional factor that limits insiders’ acquisition of private control benefits, thereby reducing their incentives to manage accounting earnings by ensuring strong and well-enforced rights for outside investors. Finally, private control benefits are the advantages that insiders can gain from controlling a firm, which can include financial gains or other benefits not shared with other stakeholders [@Leuz_2003].
3.	While pulling the data for analysis, I encountered negative values for some key financial metrics such as operating income (`item1250`) and net income before preferred dividends (`item1651`). The paper by @Leuz_2003 does not explicitly specify how to handle negative values in key financial metrics. For the purpose of this replication study, I included negative values in the analysis. This ensures that the analysis captures the full spectrum of earnings management activities across different countries.
4.	Another potential source of discrepancies between the original and replicated tables may be the choice of variables pulled from Worldscope. For example, in the `wrds_ws_funda table`, both `item1151` and `item4051` are labeled as “DEPRECIATION, DEPLETION AND AMORTIZATION” [@WRDS_WS_Funda_2024]. I chose `item1151` from the Income Statement rather than `item4051` from the Cash Flow Statement, based on the *Worldscope Balancing Model - Industrials* by Wharton Research Data Services [-@WRDS_Balancing_Model_2024]. Since the authors do not specify the choice of variables used from the database, this could cause differences in the results.
5.	The EM measures are based on scaled variables (e.g., operating cash flow scaled by lagged total assets). As such, the currency of the relevant data items should not affect the results as long as the same currency is used consistently for both the numerator and the denominator. This approach ensures comparability across different countries, regardless of their local currencies. Additionally, according to a document by @Thomson_Financial_2007 [p.20], all Worldscope data is consistently reported in the local currency of each firm’s country of domicile, eliminating the need for currency conversions in this project.

By following the steps provided in @sec-replication_steps and adhering to the assumptions made, I successfully replicated the analysis and produced the required table. A thorough step-by-step approach, with each step clearly documented, helped to understand and verify the outputs.

# Replication Steps {#sec-replication_steps}

## Step 1: Pulling the Data and Managing the Databases
Unlike Assignment I, where the data was provided externally, Assignment III involves pulling data directly from the Worldscope database, merging relevant tables, and preparing the data for further analysis from raw data to final output.

To ensure data integrity, I filtered out rows with empty `item6105` (Worldscope Permanent ID) values, as this identifier is crucial for firm/year-level filtering in the data preparation step. A total of 125 observations from the dynamic data and 10,306 observations from the static data were removed. The dynamic and static datasets were then merged on the `item6105` identifier, representing the unique Worldscope Permanent ID. According to @WRDS_Overview_2024, this identifier should be used consistently within Worldscope data, as it remains stable over time. An inner join was used for this merge to ensure that only complete and consistent data from both tables were retained.

Additionally, a filter was applied to retain only company rows. This was achieved by selecting rows where the `item6100` field equals ‘C’, indicating that the Worldscope Identifier represents a company. This step ensures that the analysis includes only company data, excluding averages, exchange rates, securities, or stock indices, as advised by @WRDS_Overview_2024.

In line with the methodology used by @Leuz_2003, financial institutions were removed from the analysis based on their Standard Industrial Classification (SIC) codes. This was done by filtering out rows where the `item7021` identifier, representing the SIC code, falls within the range of 6000 to 6999. This ensures that the dataset focuses solely on non-financial companies.

Further filtering ensured that only data from the 31 countries specified in the paper were included. These steps were applied to reduce the dataset size and improve the workflow.

Moreover, the configuration file utilized additional refinements as proposed by @WRDS_Overview_2024. Specifically, using "A" (Annual) in the `freq` variable ensured that the data represents financial information reported on an annual basis, consistent with the methodology outlined in the paper by @Leuz_2003. This excludes data reported on a current, daily, or quarterly basis. After retaining only relevant company data, filtering out financial institutions, and focusing on the specified countries, the processed data was saved to a CSV file in the `data/pulled` directory, as specified in configuration file `config/pull_data_cfg`​.

## Step 2: Data Preparation {#sec-step2_preparation}
To verify the pulled data, I checked the dataset for duplicate firm-year observations based on the combination of the Worldscope Permanent ID (`item6105`) and the fiscal year (`year_`). I confirmed that no duplicates were present, ensuring the accuracy of the data for further analysis. As the original study does not specify the net income measure, I used `item1651` (Net Income before Preferred Dividends) as the variable for net income. This choice aligns with the final net income figure reported in the income statement, based on *Worldscope Balancing Model - Industrials* by Wharton Research Data Services [-@WRDS_Balancing_Model_2024].

To prepare the data sample in line with the methodology outlined in the paper, it was essential to follow the requirements set by @Leuz_2003. First, countries with sufficient firm-year observations were filtered. Each country needed at least 300 firm-year observations for key accounting variables, including total assets, sales, net income, and operating income. In this step, no countries were eliminated, as all countries met the requirement, which aligns with the paper’s overview of countries (31 countries in total).

In the second filtration step, firms with adequate consecutive data were identified. Each firm had to have income statement and balance sheet information for at least three consecutive years, with all key accounting variables mentioned in previous step being present. If a firm had at least three consecutive years of complete data at any point, all its data entries were retained in the final dataset. Therefore, only those countries and firms that met these criteria for all specified variables were retained in the dataset.

During the preparation step, 8,265 firms and 20,521 firm-year observations (all due to the second filtration step) were dropped, resulting in a final dataset with 18,040 non-financial firms and 123,469 firm-year observations. The differences in the numbers between the prepared dataset and the figures mentioned in the paper (70,955 firm-year observations and 8,616 non-financial firms) could be due to assumptions outlined in @sec-research_design_assumptions, such as variations in initial datasets, data updates, and filtering criteria. However, the original paper might have included additional data cleaning steps not explicitly mentioned, such as handling outliers, specific industry exclusions, or other criteria, which could affect the final counts.

Given that the number of observations in this project is significantly higher than in the original study, Table 1 from @Leuz_2003 was partially replicated (focusing only on columns related to countries and firm-year observations) to highlight specific discrepancies that could arise for certain countries. So, the results in Table 2 show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset. The trend in the number of firm-year observations across countries is consistent with the original study, with larger capital markets like Japan, the United Kingdom, and the United States having the highest numbers. The distribution of observations in my dataset shows higher mean and median values compared to the paper by @Leuz_2003, indicating a greater overall number of firm-year observations, while the minimum values are relatively close, suggesting that the overall distribution of observations across countries remains broadly similar.

However, there are significant discrepancies, particularly for the United States, which has 41,585 observations compared to 3,792 in the original study. @Leuz_2003 focused only on S&P 500 companies for the U.S., while this analysis uses a more comprehensive dataset from the Worldscope database, which includes detailed financial statement data and profile data on public companies globally. This broader scope results in a higher number of firm-year observations for the United States. Despite the differences in absolute numbers, the relative trends across countries align with the original paper.

```{python}
#| label: table_1 
#| echo: false
#| output: asis

print(prep_latex_table(
    results['table_1'],
    caption='Replicated table - Descriptive statistics: Number of firm-year observations per country',
    label='tab:table_1'
))
```

## Step 3: Analysis Implementation and Table Reproduction
In this replication step, the calculation of the EM measures is implemented according to the methodology described by @Leuz_2003. The study employs four distinct proxies that capture different dimensions of earnings management, such as earnings smoothing and accrual manipulations. These measures are used to assess the pervasiveness of earnings management across countries. This step ensures that all statistical methods and groupings are accurately replicated, following the paper’s approach to capture the extent to which corporate insiders use accounting discretion to obscure their firm’s true economic performance.

### Smoothing Reported Operating Earnings Using Accruals (EM1) {#sec-EM1}
EM1 is a measure that assesses how much insiders (e.g., company executives) smooth or stabilize reported earnings by manipulating accruals. The concept is that firms may attempt to reduce fluctuations in their reported operating income to present a more stable financial performance to investors. This measure reflects the degree to which insiders use accounting discretion to smooth reported earnings, with lower values of EM1 suggesting more extensive earnings smoothing.

To calculate this measure:

1.	Compute Accruals by using changes in various components of a firm’s balance sheet and income statement, such as total current assets, cash, current liabilities, short-term debt, taxes payable, and depreciation:

$$
\begin{aligned}
\text{Accruals}_{it} &= (\Delta \text{CA}_{it} - \Delta \text{Cash}_{it}) - (\Delta \text{CL}_{it} - \Delta \text{STD}_{it} - \Delta \text{TP}_{it}) - \text{Dep}_{it} 
& \\
\text{where:} \\
\Delta \text{CA}_{it} & = \text{Change in total current assets} \\
\Delta \text{Cash}_{it} & = \text{Change in cash and cash equivalents} \\
\Delta \text{CL}_{it} & = \text{Change in total current liabilities} \\
\Delta \text{STD}_{it} & = \text{Change in short-term debt included in current liabilities} \\
\Delta \text{TP}_{it} & = \text{Change in income taxes payable} \\
\text{Dep}_{it} & = \text{Depreciation and amortization expense for firm i in year t} 
\end{aligned}
$$

2.	Derive Operating Cash Flow (CFO) that is indirectly derived from operating income minus accruals:
$$
\text{CFO}_{it} = \text{Operating Income}_{it} - \text{Accruals}_{it}
$$

3.	Calculate EM1, which is determined as the country’s median ratio of the firm-level standard deviations of operating income and operating cash flow. Both standard deviations are scaled by lagged total assets to control for differences in firm size:
$$
\text{EM1} = \text{Median}\left(\frac{\frac{\sigma(\text{Operating Income}_{it})}{\text{Total Assets}_{i(t-1)}}}{\frac{\sigma(\text{Operating Cash Flow}_{it})}{\text{Total Assets}_{i(t-1)}}}\right)
$$

### Smoothing and the Correlation Between Changes in Accounting Accruals and Operating Cash Flows (EM2)
EM2 is designed to capture the extent to which insiders use accounting discretion to smooth reported earnings by examining the correlation between changes in accruals and changes in operating cash flows. A stronger negative correlation between these variables indicates that accruals are being used to buffer the impact of cash flow fluctuations, potentially masking the true economic performance of the firm.

To calculate this measure:

1. Compute Accruals and CFO using same Formulas used in @sec-EM1.

2. Calculate the change in Accruals and CFO for each firm-year:
$$
\Delta \text{Accruals}_{it} = \text{Accruals}_{it} - \text{Accruals}_{i(t-1)}
$$
$$
\Delta \text{CFO}_{it} = \text{CFO}_{it} - \text{CFO}_{i(t-1)}
$$

3. Scale both changes by lagged total assets:
$$
\frac{\Delta \text{Accruals}_{it}}{\text{Total Assets}_{i(t-1)}}, \quad \frac{\Delta \text{CFO}_{it}}{\text{Total Assets}_{i(t-1)}}
$$

4. Calculate EM2, which is computed as the Spearman correlation between these scaled changes, aggregated at the country level over the pooled set of firms within each country, meaning that the correlation reflects the aggregated relationship between changes in accruals and changes in CFO across all firms in that country:
$$
\text{EM2} = \rho\left(\frac{\Delta \text{Accruals}_{it}}{\text{Total Assets}_{i(t-1)}}, \frac{\Delta \text{CFO}_{it}}{\text{Total Assets}_{i(t-1)}}\right)
$$

### Discretion in Reported Earnings: The Magnitude of Accruals (EM3)
EM3 is a measure that obtains the degree of discretion that insiders exercise when reporting earnings, particularly when it comes to misstating a firm’s economic performance. It focuses on the magnitude of accruals as a proxy for earnings management. Insiders can use accruals to overstate or understate earnings, thereby achieving certain financial targets or presenting a skewed view of the firm’s performance. Higher scores of EM3 indicate more earnings management, as larger magnitudes of accruals relative to cash flow from operations suggest greater discretion in financial reporting.

To calculate this measure:

1. Compute Accruals and CFO using same Formulas used in @sec-EM1.

2. Calculate the absolute values of accruals and CFO for each firm-year.

3. EM3 is calculated as the country-level median ratio of the absolute value of firms’ accruals to the absolute value of their cash flow from operations. This ratio is used as it scales the magnitude of accruals to a measure of the firm’s size and performance, allowing for comparison across firms of different sizes and performance levels:
$$
\text{EM3} = \text{Median}\left(\frac{|\text{Accruals}_{it}|}{|\text{CFO}_{it}|}\right)
$$

### Discretion in Reported Earnings: Small Loss Avoidance (EM4)
EM4 captures the extent to which firms manage their reported earnings to avoid small losses. This is an important measure that quantifies the propensity of firms to avoid small losses through earnings management. By calculating the ratio of small profits to small losses, it provides a clear indication of how widespread this practice is within a particular country. A higher EM4 value suggests that firms in a given country are more likely to use accounting discretion to convert small losses into small profits, reflecting a tendency toward earnings management. Note that I used `item1651` (Net Income before Preferred Dividends) as the variable for net earnings, as mentioned in @sec-step2_preparation. I also ensured that each country included in the EM4 calculation had at least five small loss observations, as per the methodology outlined by @Leuz_2003. The review found that no countries were excluded from the analysis due to this criterion.

EM4 is calculated as the ratio of the number of small profits to the number of small losses for each country, where small profits and small losses are determined by scaling net earnings by lagged total assets:
$$
\text{EM4} = \frac{\text{Small Profits}}{\text{Small Losses}}
$$
$$
\begin{alignedat}{2}
\text{where:} \\ \quad & \text{Small Profits} && = \frac{\text{Net Earnings}_{it}}{\text{Total Assets}_{it-1}} \in [0, 0.01] \\
& \text{Small Losses} && = \frac{\text{Net Earnings}_{it}}{\text{Total Assets}_{it-1}} \in [-0.01, 0)
\end{alignedat}
$$

### Aggregate Measure of Earnings Management
The aggregate earnings management score is a summary measure that captures the overall level of earnings management in each country. This measure is calculated by ranking countries on each of the four individual earnings management measures (EM1-EM4), where a higher rank indicates a greater degree of earnings management. The final aggregate score is the average of these ranks, providing a comprehensive view of a country’s earnings management practices. The countries are sorted in descending order based on their aggregate score.

To calculate this measure:
$$
\text{Aggregate Earnings Management Score}_{\text{country}} = \frac{\text{Rank}_{\text{EM1}} + \text{Rank}_{\text{EM2}} + \text{Rank}_{\text{EM3}} + \text{Rank}_{\text{EM4}}}{4}
$$

# Results {#sec-results}
The comparison between the calculated EM1 values and those reported in the original paper reveals a strong alignment, with only minor differences observed across most countries. The overall pattern of EM1 values is consistent, effectively capturing the underlying trends in earnings management across different countries. Any differences, such as those for the Philippines, Taiwan, and Norway, are relatively small and likely due to variations in data sources, updates, or slight methodological differences, as outlined in @sec-research_design_assumptions. The summary statistics (mean, median, min, max, and standard deviation) also closely match those in the original study, further supporting the robustness of the replication process. The results indicate that Anglo-American countries tend to have higher EM1 values, suggesting less earnings smoothing compared to Continental Europe and Asia, consistent with the conclusion of @Leuz_2003 that earnings are smoother in Continental Europe and Asia than in Anglo-American countries after controlling for the volatility of cash flows.

The replication findings regarding EM2 indicate that earnings smoothing is more pronounced in countries like Greece and Japan, while countries like Canada and the United States exhibit less pervasive smoothing behavior, in line with the findings of @Leuz_2003. Despite some differences in magnitude, the overall trends are consistent, reinforcing the idea that earnings smoothing varies significantly across different countries. In fact, these small variations do not significantly alter the overall conclusions. Specifically, the minimum and maximum EM2 values in the replication closely match the original findings, with Greece showing the most earnings smoothing (min: -0.942 vs. -0.928) and Norway the least (max: -0.786 vs. -0.722). This consistency reinforces the accuracy of the replication.

The replicated EM3 results are very similar to those of the original study, with close alignment in summary statistics and consistent identification of South Africa and Germany as the countries with the minimum and maximum EM3 values. Additionally, the replication confirms that the magnitude of firms' accruals relative to their operating cash flows is smaller in the U.K. and the U.S. compared to Austria, Germany, and South Korea, consistent with the original findings.

The replicated results show significant regional differences in EM4 values, reflecting varying degrees of loss avoidance. Spain has the highest EM4 value, highlighting strong loss avoidance among European firms. Norway, with the lowest EM4 value, corresponds to the minimum value in the original paper, confirming less aggressive loss avoidance. High EM4 values in Asian countries like Indonesia (5.667) and India (6.000) further support the trend that European and Asian firms exhibit greater loss avoidance than their Anglo-American counterparts, such as the United States (1.912) and Canada (2.402), which have lower EM4 values. These results align with the original study's conclusions on regional loss avoidance behavior.

The comparison between the original and replicated aggregate earnings management scores reveals a strong overall alignment with consistent key trends. In the original study, Austria and Greece are at the top of the rankings with aggregate scores of 28.3, indicating the highest levels of earnings management. In the replication, Greece and Austria remain at the top with slightly adjusted scores of 27.5 and 27.2, respectively. Taiwan and South Korea also rank high, mirroring their positions in the original study. While some differences in specific rankings and scores exist, such as the United States consistently having the lowest level of earnings management (2.0 in the original and 3.5 in the replication), this slight variation could be attributed to differences in the datasets, particularly due to the use of a more comprehensive dataset from the Worldscope database, which includes financial data on U.S. public companies globally, in contrast to only S&P index firms used in the original paper, as discussed in @sec-step2_preparation. In general, these variations are minor and do not undermine the overall consistency.

In summary, the replication effectively captures the original study's key findings, confirming that European and Asian countries generally exhibit higher levels of earnings management than their Anglo-American counterparts. The strong similarity between the original and replicated results underscores the robustness and reliability of the findings.

# Conclusion {#sec-conclusion}
The project effectively demonstrates the use of an open science and collaborative workflow for analyzing earnings management measures across various countries. By following a systematic approach and utilizing the TRR 266 Template for Reproducible Empirical Accounting Research, I successfully replicated an empirical table from @Leuz_2003 and provided insights into trends in earnings management practices, which, similar to the original study, vary significantly across countries.

@Leuz_2003 document systematic differences in the level of earnings management across 31 countries, suggesting that insiders use earnings management to conceal firm performance and protect their private control benefits. The study further argues that stronger investor protection reduces earnings management by limiting insiders’ ability to secure private benefits. By conducting a descriptive cluster analysis following the descriptive statistics replicated in this project, the paper identifies groupings of countries with similar institutional characteristics, demonstrating that earnings management varies systematically across these institutional clusters.

The replicated results reveal significant differences in earnings management practices across countries, with consistent patterns in the relative magnitudes of the individual measures. These differences are particularly notable in areas such as loss avoidance and earnings smoothing. In line with the original study, the replication suggests that firms in countries with more developed equity markets, dispersed ownership structures, strong investor rights, and robust legal enforcement tend to engage in less earnings management. This consistent finding across both the original and replicated studies underscores the critical role of the institutional environment in shaping corporate earnings management behaviors on an international scale.

The replication analysis also reveals that the individual earnings management measures are highly correlated, resulting in similar rankings across the four measures and the aggregate earnings management score. This consistency supports the original study’s conclusion that a single underlying factor likely represents the four individual measures. As in the original results, the replicated country rankings based on the aggregate score show high ranks for countries like Greece, Austria, and South Korea, and lower ranks for countries such as Australia, the U.K., and the U.S. This reinforces the robustness of the aggregate measure in capturing the overall level of earnings management across different countries.

This final assignment required the comprehensive application of all the skills learned and feedback received from previous assignments, making it a more complex project workflow. In the future, this repository can be cloned or forked (if made public) to kickstart further projects on earnings management measures analysis. Additionally, the included presentation template can be rendered alongside the paper for further academic use, providing a visually engaging summary of the research findings. Thanks for reading!

\pagebreak

\setcounter{table}{0}
\renewcommand{\thetable}{\arabic{table}}

# References {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent

### 2024-11-12 doc/presentation.qmd
Initial commit

---
title: |
  | Earnings Management and Investor Protection:
  | Accounting Reading Group - Assignment III
author: "Melisa Mazaeva"
institute: "Humboldt-Universität zu Berlin – Summer 2024"
date: today
fontsize: "9pt"
pdf-engine: xelatex # pdflatex creates rastered fonts
format: 
  beamer:
    slide-level: 3
    number-sections: true
    toc: false
header-includes:
- \usepackage{booktabs} 
- \usepackage{threeparttable}
- \usepackage{graphicx}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\fontsize{7}{7}\selectfont}  # This sets the font size for verbatim environments
- \input{beamer_theme_trr266.sty}
bibliography: references.bib
biblio-style: apsr
---

```{python}
#| echo: false
#| output: false

import pickle
import pandas as pd

# Load the results from the existing pickle file for table_1
with open('../data/generated/table_1.pickle', 'rb') as f:
    results = pickle.load(f)

# Load the results from the new pickle file for final_table and summary_stats
with open('../output/em_results.pickle', 'rb') as f:
    em_results = pickle.load(f)

def escape_special_characters(df):
    df.columns = [col.replace('_', '\\_').replace('#', '\\#') for col in df.columns]
    return df

# Custom formatter function
def custom_float_format(x):
    return f"{x:.3f}".lstrip('0') if pd.notnull(x) else ""

def prep_latex_table(df, caption=None, label=None, fontsize='\\fontsize{11}{13}\\selectfont', description=None):
    df = escape_special_characters(df.reset_index(drop=True))  # Drop the index column from output tables for better readability
    num_columns = len(df.columns)
    column_format = 'l' + 'r' * (num_columns - 1)

    latex_table = df.to_latex(
        column_format=column_format,
        index=False, 
        float_format=custom_float_format,
        escape=False  # Important to disable escaping so LaTeX commands in column names are respected
    )

    # Construct description to the table
    full_caption = f"{caption}\\\\\n\\textnormal{{{description}}}" if description else caption

    latex_table_lines = [
        "\\begin{table}[htbp]",
        "\\centering",
        fontsize,  # Set font size for the table
        latex_table,
        f"\\caption{{{full_caption}}}" if full_caption else "",
        f"\\label{{{label}}}" if label else "",
        "\\end{table}"
    ]
    return "\n".join(line for line in latex_table_lines if line)
```

# Motivation

## Understanding Systematic Differences in Earnings Management across Countries

- The paper of @Leuz_2003 focuses on the relationship between investor protection and earnings management.
- Investor protection plays a crucial role in limiting the extent of earnings management by reducing the opportunities for managers to manipulate earnings.
- Analyzing the relationship between EM and investor protection across different countries provides insights into how legal and institutional frameworks impact financial reporting practices.
- Earnings management (EM) measures reflect the extent to which managers manipulate financial reports to meet certain objectives, such as meeting earnings targets or influencing stock prices.

## Project Objective

- The primary objective is to demonstrate a reproducible and collaborative research workflow by calculating and analyzing EM measures across different countries.
- The study aims to replicate and extend the findings of @Leuz_2003, exploring the relationship between EM and investor protection using Worldscope datasets for the time range 1990 to 1999.
- The analysis provides insights into the levels of earnings management across countries, comparing them with the original findings of Leuz, Nanda, and Wysocki (2003).

## Project Relevance

- Understanding EM trends across countries aids in assessing global financial transparency.
- Promotes transparency and reproducibility in empirical accounting research.

# Research Design

- Data sourced from the Worldscope Database (1990-1999) July 2024 version.
- Replication includes calculation of **four earnings management measures**:
  1. Earnings smoothing using accruals (EM1)
  2. Correlation between changes in accruals and cash flows (EM2)
  3. Magnitude of accruals (EM3)
  4. Ratio of small profits to small losses (EM4)

# Assumptions

- **Database Version**: Analysis uses the July 2024 Worldscope data, which may differ from the November 2000 version due to updates and restatements.
- **Key Terms**: Earnings management (EM) and investor protection defined as per @Leuz_2003 for consistency.
- **Handling Negative Values**: Negative values in key metrics (e.g., operating income) were included to capture the full scope of EM activities.
- **Variable Selection**: Chose specific Worldscope variables (e.g., item1151 over item4051) based on best practices; differences may affect results.
- **Currency Consistency**: EM measures are based on scaled variables; currency consistency ensures comparability across countries.

# Data - Replication Steps

Step 1: Pulling the Data and Managing the Databases

Step 2: Data Preparation

+  To verify the robustness of descriptive statistics before the main replication, Table 1 from Leuz et al. was replicated, yielding results that closely align with the original study:
---
```{python}
#| label: table_1 
#| echo: false
#| output: asis

print(prep_latex_table(
    results['table_1'],
    caption='Replicated table - Descriptive statistics: Number of firm-year observations per country',
    label='tab:table_1',
    fontsize='\\fontsize{5pt}{6pt}\\selectfont'  # Set smaller font size
))
```
---

Step 3: Analysis Implementation and Table Reproduction

- Smoothing Reported Operating Earnings Using Accruals (EM1)
    $$
    \text{EM1} = \text{Median}\left(\frac{\frac{\sigma(\text{Operating Income}_{it})}{\text{Total Assets}_{i(t-1)}}}{\frac{\sigma(\text{Operating Cash Flow}_{it})}{\text{Total Assets}_{i(t-1)}}}\right)
    $$
- Smoothing and the Correlation Between Changes in Accounting Accruals and Operating Cash Flows (EM2)
    $$
    \text{EM2} = \rho\left(\frac{\Delta \text{Accruals}_{it}}{\text{Total Assets}_{i(t-1)}}, \frac{\Delta \text{CFO}_{it}}{\text{Total Assets}_{i(t-1)}}\right)
    $$
- Discretion in Reported Earnings: The Magnitude of Accruals (EM3)
    $$
    \text{EM3} = \text{Median}\left(\frac{|\text{Accruals}_{it}|}{|\text{CFO}_{it}|}\right)
    $$

 ---
- Discretion in Reported Earnings: Small Loss Avoidance (EM4)
    $$
    \text{EM4} = \frac{\text{Small Profits}}{\text{Small Losses}}
    $$
 - Aggregate Measure of Earnings Management
    $$
    \text{Aggregate EM Score}_{\text{country}} = \frac{\text{Rank}_{\text{EM1}} + \text{Rank}_{\text{EM2}} + \text{Rank}_{\text{EM3}} + \text{Rank}_{\text{EM4}}}{4}
    $$

# Results

```{python}
#| label: final_combined_table
#| echo: false
#| output: asis

# Rename the columns to ensure consistency with paper
em_results['final_combined_table'].rename(columns={
    'item6026': 'Country',
    'EM1': 'EM1 (−)',
    'EM2': 'EM2 (−)',
    'EM3': 'EM3 (+)',
    'EM4': 'EM4 (+)',
    'Aggregate_EM_Score': 'Aggregate EM score'
}, inplace=True)

print(prep_latex_table(
    em_results['final_combined_table'], 
    caption='Replicated table - Country scores for earnings management measures (Sorted by aggregate earnings management)',
    label='tab:final_combined_table',
    fontsize='\\fontsize{4pt}{5pt}\\selectfont',  # Set smaller font size

))
```

# Conclusion

## Summary of Findings
- EM1 Alignment: Strong alignment with minor differences, reflecting consistent trends in earnings management across countries.
- EM2 Findings: Similar patterns observed; Greece and Japan show more smoothing, while Canada and the U.S. show less.
- EM3 Consistency: Replicated results closely match the original, confirming trends in accrual magnitudes.
- EM4 Regional Differences: Significant regional variations in loss avoidance, consistent with original findings.
- Aggregate Scores: Strong alignment between original and replicated scores, confirming key trends across countries.
- **Overall Conclusion**: Replication confirms higher EM in European and Asian countries compared to Anglo-American counterparts.

## Key Insights

- Significant differences in earnings management practices across countries.
- Strong alignment with original findings in Leuz, Nanda, and Wysocki (2003).
- The project structure demonstrated the effectiveness of an open science and collaborative workflow.
- **Future Work**: Repository can be cloned or forked for further analysis.

# References {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent

### 2024-11-12 code/python/prepare_data.py

@@ -1,0 +1,110 @@
+# --- Header -------------------------------------------------------------------
+# Prepare the pulled data for further analysis as per the requirements by Leuz et al.
+#
+# (C) Melisa Mazaeva -  See LICENSE file for details
+# ------------------------------------------------------------------------------
+
+import pandas as pd
+import pickle
+from utils import read_config, setup_logging
+
+log = setup_logging()
+def main():
+    log.info("Preparing data for analysis ...")
+    cfg = read_config('config/prepare_data_cfg.yaml')
+
+    # Load the pulled data
+    wrds_data = pd.read_csv(cfg['worldscope_sample_save_path'])
+    initial_obs_count_pulled = len(wrds_data)
+    initial_firm_count_pulled = len(wrds_data['item6105'].unique())
+    log.info(f"Initial number of observations after pulling data: {initial_obs_count_pulled}")
+    log.info(f"Initial number of firms after pulling data: {initial_firm_count_pulled}")
+
+    # Check for duplicate firm-year observations
+    dup_obs = wrds_data[wrds_data.duplicated(subset=['item6105', 'year_'], keep=False)]
+    if not dup_obs.empty:
+        log.warning(f"Found {dup_obs.shape[0]} duplicate firm-year observations. Removing duplicates.")
+        wrds_data = wrds_data.drop_duplicates(subset=['item6105', 'year_'], keep='first')
+
+    # Filter countries with at least 300 firm-year observations for key accounting variables as in paper
+    filtered_countries_data, eliminated_countries = filter_countries(wrds_data)
+
+    # Print eliminated countries
+    if eliminated_countries:
+        log.info(f"Countries eliminated after filtration: {', '.join(eliminated_countries)}")
+    else:
+        log.info("No countries were eliminated after filtration.")
+
+    # Filter firms with at least three consecutive years of income statement and balance sheet information as in paper
+    initial_firm_count = len(filtered_countries_data['item6105'].unique())
+    initial_obs_count = len(filtered_countries_data)
+    
+    filtered_firms_data = filter_firms(filtered_countries_data)
+
+    final_firm_count = len(filtered_firms_data['item6105'].unique())
+    final_obs_count = len(filtered_firms_data)
+
+    firms_dropped = initial_firm_count - final_firm_count
+    obs_dropped = initial_obs_count - final_obs_count
+
[…]

### 2024-11-12 code/python/pull_wrds_data.py

@@ -1,0 +1,98 @@
+# --- Header -------------------------------------------------------------------
+# See LICENSE file for details
+#
+# This code pulls data from WRDS Worldscope Database
+# ------------------------------------------------------------------------------
+import os
+from getpass import getpass
+import dotenv
+
+import pandas as pd
+from utils import read_config, setup_logging
+import wrds
+
+log = setup_logging()
+
+def main():
+    '''
+    Main function to pull data from WRDS.
+
+    This function reads the configuration file, gets the WRDS login credentials, and pulls the data from WRDS.
+
+    The data is then saved to a csv file.
+    '''
+    cfg = read_config('config/pull_data_cfg.yaml')
+    wrds_login = get_wrds_login()
+    wrds_data = pull_wrds_data(cfg, wrds_login)
+    wrds_data.to_csv(cfg['worldscope_sample_save_path'], index=False)
+
+
+def get_wrds_login():
+    '''
+    Gets the WRDS login credentials.
+    '''
+    if os.path.exists('secrets.env'):
+        dotenv.load_dotenv('secrets.env')
+        wrds_username = os.getenv('WRDS_USERNAME')
+        wrds_password = os.getenv('WRDS_PASSWORD')
+        return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
+    else:
+        wrds_username = input('Please provide a wrds username: ')
+        wrds_password = getpass(
+            'Please provide a wrds password (it will not show as you type): ')
+        return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
+
+def pull_wrds_data(cfg, wrds_authentication):
+    '''
+    Pulls WRDS access data.
+    '''
+    db = wrds.Connection(
+        wrds_username=wrds_authentication['wrds_username'], wrds_password=wrds_authentication['wrds_password']
+    )
+
+    log.info('Logged on to WRDS ...')
+
+    dyn_var_str = ', '.join(cfg['dyn_vars'])
+    stat_var_str = ', '.join(cfg['stat_vars'])
+
+    log.info("Pulling dynamic Worldscope data ... ")
+    wrds_data_dynamic = db.raw_sql(
+        f"SELECT {dyn_var_str} FROM tr_worldscope.wrds_ws_funda WHERE {cfg['cs_filter']}"
+    )
+    log.info("Pulling dynamic Worldscope data ... Done!")
+
+    log.info("Pulling static Worldscope data ... ")
[…]

### 2024-11-15 code/python/pull_wrds_data old.py
config data done

@@ -1,7 +1,7 @@
 # --- Header -------------------------------------------------------------------
 # See LICENSE file for details
 #
-# This code pulls data from WRDS Audit Analytics Database
+# This code pulls data from WRDS Audit Analytics Database 
 # ------------------------------------------------------------------------------
 import os
 from getpass import getpass

### 2024-11-16 code/python/pull_wrds_data.py
pull data

@@ -1,8 +1,9 @@
 # --- Header -------------------------------------------------------------------
 # See LICENSE file for details
 #
-# This code pulls data from WRDS Audit Analytics Database
+# This code pulls data from WRDS Audit Europe Database for Figure 3 replication
 # ------------------------------------------------------------------------------
+
 import os
 from getpass import getpass
 import dotenv
@@ -24,7 +25,7 @@ def main():
     cfg = read_config('config/pull_data_cfg.yaml')
     wrds_login = get_wrds_login()
     wrds_data = pull_wrds_data(cfg, wrds_login)
-    wrds_data.to_csv(cfg['worldscope_sample_save_path'], index=False)
+    wrds_data.to_csv(cfg['output_save_path'], index=False)
 
 
 def get_wrds_login():
@@ -37,14 +38,14 @@ def get_wrds_login():
         wrds_password = os.getenv('WRDS_PASSWORD')
         return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
     else:
-        wrds_username = input('Please provide a wrds username: ')
+        wrds_username = input('Please provide a WRDS username: ')
         wrds_password = getpass(
-            'Please provide a wrds password (it will not show as you type): ')
+            'Please provide a WRDS password (it will not show as you type): ')
         return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
 
 def pull_wrds_data(cfg, wrds_authentication):
     '''
-    Pulls WRDS access data.
+    Pulls WRDS Audit Europe Transparency Report data.
     '''
     db = wrds.Connection(
         wrds_username=wrds_authentication['wrds_username'], wrds_password=wrds_authentication['wrds_password']
@@ -52,45 +53,24 @@ def pull_wrds_data(cfg, wrds_authentication):
 
     log.info('Logged on to WRDS ...')
 
-    dyn_var_str = ', '.join(cfg['dyn_vars'])
-    stat_var_str = ', '.join(cfg['stat_vars'])
+    # Select only the required variables for Transparency Report
+    selected_vars_str = ', '.join(cfg['selected_vars'])
+    
+    # Apply filter for 2021 and specific countries
+    country_filter = ', '.join([f"'{country}'" for country in cfg['included_countries']])
+    query = f"""
+        SELECT {selected_vars_str}
+        FROM audit_europe.feed76_transparency_reports
+        WHERE REPORT_YEAR = 2021
[…]

### 2024-11-16 code/python/pull_wrds_data.py
fresh readme tip

@@ -1,9 +1,8 @@
 # --- Header -------------------------------------------------------------------
 # See LICENSE file for details
 #
-# This code pulls data from WRDS Audit Europe Database for Figure 3 replication
+# This code pulls data from WRDS Audit Analytics Database 
 # ------------------------------------------------------------------------------
-
 import os
 from getpass import getpass
 import dotenv

### 2024-11-16 code/python/pull_wrds_data.py
pull data final

@@ -24,7 +24,7 @@ def main():
     cfg = read_config('config/pull_data_cfg.yaml')
     wrds_login = get_wrds_login()
     wrds_data = pull_wrds_data(cfg, wrds_login)
-    wrds_data.to_csv(cfg['output_save_path'], index=False)
+    wrds_data.to_csv(cfg['audit_analytics_save_path'], index=False)
 
 
 def get_wrds_login():
@@ -52,9 +52,8 @@ def pull_wrds_data(cfg, wrds_authentication):
 
     log.info('Logged on to WRDS ...')
 
-    # Select only the required variables for Transparency Report
+    # Select only the required variables from Transparency Reports
     selected_vars_str = ', '.join(cfg['selected_vars'])
-    
     # Apply filter for 2021 and specific countries
     country_filter = ', '.join([f"'{country}'" for country in cfg['included_countries']])
     query = f"""

### 2024-11-16 code/python/prepare_data old.py
prepare data intermediary

@@ -1,5 +1,5 @@
 # --- Header -------------------------------------------------------------------
-# Prepare the pulled data for further analysis as per the requirements by EC Report (2024)
+# Prepare the pulled data for further analysis as per the requirements by Leuz et al.
 #
 # (C) Melisa Mazaeva -  See LICENSE file for details
 # ------------------------------------------------------------------------------

### 2024-11-16 code/python/prepare_data.py
prep

@@ -9,102 +9,58 @@ import pickle
 from utils import read_config, setup_logging
 
 log = setup_logging()
+
 def main():
     log.info("Preparing data for analysis ...")
     cfg = read_config('config/prepare_data_cfg.yaml')
 
     # Load the pulled data
-    wrds_data = pd.read_csv(cfg['worldscope_sample_save_path'])
-    initial_obs_count_pulled = len(wrds_data)
-    initial_firm_count_pulled = len(wrds_data['item6105'].unique())
-    log.info(f"Initial number of observations after pulling data: {initial_obs_count_pulled}")
-    log.info(f"Initial number of firms after pulling data: {initial_firm_count_pulled}")
+    transparency_data = pd.read_csv(cfg['audit_analytics_save_path'])
+    initial_obs_count = len(transparency_data)
+    log.info(f"Initial number of observations after pulling data: {initial_obs_count}")
 
-    # Check for duplicate firm-year observations
-    dup_obs = wrds_data[wrds_data.duplicated(subset=['item6105', 'year_'], keep=False)]
-    if not dup_obs.empty:
-        log.warning(f"Found {dup_obs.shape[0]} duplicate firm-year observations. Removing duplicates.")
-        wrds_data = wrds_data.drop_duplicates(subset=['item6105', 'year_'], keep='first')
+    # Handle missing Auditor Network values
+    transparency_data['network_flag'] = transparency_data['auditor_network'].fillna('Unaffiliated')
+    missing_networks = transparency_data['network_flag'].value_counts().get('Unaffiliated', 0)
+    log.info(f"Number of observations with missing Auditor Network: {missing_networks}")
 
-    # Filter countries with at least 300 firm-year observations for key accounting variables as in paper
-    filtered_countries_data, eliminated_countries = filter_countries(wrds_data)
+    # Group Auditor Network to standardize values (Big 4, 10KAP, and others)
+    transparency_data = standardize_auditor_networks(transparency_data)
 
-    # Print eliminated countries
-    if eliminated_countries:
-        log.info(f"Countries eliminated after filtration: {', '.join(eliminated_countries)}")
-    else:
-        log.info("No countries were eliminated after filtration.")
+    # Save the prepared dataset
+    transparency_data.to_csv(cfg['prepared_data_save_path'], index=False)
+    log.info(f"Prepared data saved to {cfg['prepared_data_save_path']}")
 
[…]

### 2024-11-16 code/python/prepare_data.py
prep 2

@@ -1,7 +1,7 @@
 # --- Header -------------------------------------------------------------------
 # Prepare the pulled data for further analysis as per the requirements by EC Report (2024)
 #
-# (C) Melisa Mazaeva -  See LICENSE file for details
+# (C) Melisa Mazaeva - See LICENSE file for details
 # ------------------------------------------------------------------------------
 
 import pandas as pd
@@ -20,20 +20,20 @@ def main():
     log.info(f"Initial number of observations after pulling data: {initial_obs_count}")
 
     # Handle missing Auditor Network values
-    transparency_data['network_flag'] = transparency_data['auditor_network'].fillna('Unaffiliated')
-    missing_networks = transparency_data['network_flag'].value_counts().get('Unaffiliated', 0)
+    transparency_data['auditor_network'] = transparency_data['auditor_network'].fillna('Other (Blank)')
+    missing_networks = transparency_data['auditor_network'].value_counts().get('Other (Blank)', 0)
     log.info(f"Number of observations with missing Auditor Network: {missing_networks}")
 
-    # Group Auditor Network to standardize values (Big 4, 10KAP, and others)
-    transparency_data = standardize_auditor_networks(transparency_data)
+    # Map auditor networks into groups
+    transparency_data = map_auditor_networks(transparency_data)
 
     # Save the prepared dataset
     transparency_data.to_csv(cfg['prepared_data_save_path'], index=False)
     log.info(f"Prepared data saved to {cfg['prepared_data_save_path']}")
 
     # Generate summary table for Auditor Networks
-    summary_table = transparency_data.groupby('auditor_network').size().reset_index(name='# Audits')
-    summary_table.columns = ['Auditor Network', '# Audits']
+    summary_table = transparency_data.groupby('network_group').size().reset_index(name='# Audits')
+    summary_table.columns = ['Network Group', '# Audits']
     summary_table = summary_table.sort_values(by='# Audits', ascending=False)
 
     # Save summary table
@@ -43,23 +43,40 @@ def main():
 
     log.info("Preparing data for analysis ... Done!")
 
-def standardize_auditor_networks(df):
+def map_auditor_networks(df):
     """
-    Standardize Auditor Network values for grouping.
+    Map auditor networks into Big Four, 10KAP (includes Big Four firms), Unaffiliated (not subset of BIG 4 and 10KAP), and Other (Blank).
     """
[…]

### 2024-11-16 code/python/prepare_data.py
grouping audit firms succesful

@@ -5,11 +5,9 @@
 # ------------------------------------------------------------------------------
 
 import pandas as pd
-import pickle
 from utils import read_config, setup_logging
 
 log = setup_logging()
-
 def main():
     log.info("Preparing data for analysis ...")
     cfg = read_config('config/prepare_data_cfg.yaml')
@@ -30,16 +28,6 @@ def main():
     # Save the prepared dataset
     transparency_data.to_csv(cfg['prepared_data_save_path'], index=False)
     log.info(f"Prepared data saved to {cfg['prepared_data_save_path']}")
-
-    # Generate summary table for Auditor Networks
-    summary_table = transparency_data.groupby('network_group').size().reset_index(name='# Audits')
-    summary_table.columns = ['Network Group', '# Audits']
-    summary_table = summary_table.sort_values(by='# Audits', ascending=False)
-
-    # Save summary table
-    with open(cfg['table_1_save_path'], 'wb') as f:
-        pickle.dump({"summary_table": summary_table}, f)
-    log.info(f"Summary table saved to {cfg['table_1_save_path']}")
 
     log.info("Preparing data for analysis ... Done!")
 

### 2024-11-16 code/python/prepare_data.py
Standardize duplicate auditor network

@@ -8,6 +8,7 @@ import pandas as pd
 from utils import read_config, setup_logging
 
 log = setup_logging()
+
 def main():
     log.info("Preparing data for analysis ...")
     cfg = read_config('config/prepare_data_cfg.yaml')
@@ -22,6 +23,9 @@ def main():
     missing_networks = transparency_data['auditor_network'].value_counts().get('Other (Blank)', 0)
     log.info(f"Number of observations with missing Auditor Network: {missing_networks}")
 
+    # Map duplicate auditor networks to standardized names
+    transparency_data = standardize_auditor_network_names(transparency_data)
+
     # Map auditor networks into groups
     transparency_data = map_auditor_networks(transparency_data)
 
@@ -31,9 +35,22 @@ def main():
 
     log.info("Preparing data for analysis ... Done!")
 
+def standardize_auditor_network_names(df):
+    """
+    Standardize duplicate auditor network names to a single name.
+    """
+    duplicate_map = {
+        '|Mazars Worldwide|Praxity Global Alliance|': '|Mazars Worldwide|',
+        '|Praxity Global Alliance|TALENZ International|': '|Praxity Global Alliance|',
+        '|Morison KSi|': '|Morison International|'
+    }
+    df['auditor_network'] = df['auditor_network'].replace(duplicate_map)
+    log.info(f"Auditor Network names standardized: {duplicate_map}")
+    return df
+
 def map_auditor_networks(df):
     """
-    Map auditor networks into Big Four, 10KAP (includes Big Four firms), Unaffiliated (not subset of BIG 4 and 10KAP), and Other (Blank).
+    Group auditor networks into Big Four, 10KAP (includes Big Four firms), Unaffiliated (not subset of BIG 4 and 10KAP), and Other (Blank).
     """
     # Define Big Four firms
     big4 = [

### 2024-11-16 code/python/prepare_data.py
handle blank values

@@ -18,6 +18,10 @@ def main():
     initial_obs_count = len(transparency_data)
     log.info(f"Initial number of observations after pulling data: {initial_obs_count}")
 
+    # Check and handle blank values in vital fields
+    vital_fields = ['transparency_report_fkey', 'entity_map_fkey', 'auditor_fkey', 'trans_report_auditor_state']
+    transparency_data = check_and_handle_blanks(transparency_data, vital_fields)
+
     # Handle missing Auditor Network values
     transparency_data['auditor_network'] = transparency_data['auditor_network'].fillna('Other (Blank)')
     missing_networks = transparency_data['auditor_network'].value_counts().get('Other (Blank)', 0)
@@ -34,6 +38,20 @@ def main():
     log.info(f"Prepared data saved to {cfg['prepared_data_save_path']}")
 
     log.info("Preparing data for analysis ... Done!")
+
+def check_and_handle_blanks(df, fields):
+    """
+    Check for blank values in specified fields and remove rows with blanks.
+    """
+    initial_count = len(df)
+    for field in fields:
+        blank_count = df[field].isnull().sum()
+        if blank_count > 0:
+            log.warning(f"Found {blank_count} blank values in '{field}'. These rows will be dropped.")
+            df = df[df[field].notnull()]
+    final_count = len(df)
+    log.info(f"Rows removed due to blank values: {initial_count - final_count}")
+    return df
 
 def standardize_auditor_network_names(df):
     """

### 2024-11-16 code/python/prepare_data.py
check for joint audits

@@ -22,6 +22,9 @@ def main():
     vital_fields = ['transparency_report_fkey', 'entity_map_fkey', 'auditor_fkey', 'trans_report_auditor_state']
     transparency_data = check_and_handle_blanks(transparency_data, vital_fields)
 
+    # Check for duplicates in entity_map_fkey
+    check_for_joint_audits(transparency_data)
+
     # Handle missing Auditor Network values
     transparency_data['auditor_network'] = transparency_data['auditor_network'].fillna('Other (Blank)')
     missing_networks = transparency_data['auditor_network'].value_counts().get('Other (Blank)', 0)
@@ -52,6 +55,16 @@ def check_and_handle_blanks(df, fields):
     final_count = len(df)
     log.info(f"Rows removed due to blank values: {initial_count - final_count}")
     return df
+
+def check_for_joint_audits(df):
+    """
+    Check for duplicates in entity_map_fkey to identify potential joint audits.
+    """
+    duplicate_count = df['entity_map_fkey'].duplicated().sum()
+    if duplicate_count > 0:
+        log.warning(f"Potential joint audits detected: {duplicate_count} entities appear in multiple transparency reports.")
+    else:
+        log.info("No potential joint audits detected. Each entity is unique in the dataset.")
 
 def standardize_auditor_network_names(df):
     """

### 2024-11-16 code/python/prepare_data.py
filter_disclosed_pies

@@ -8,7 +8,6 @@ import pandas as pd
 from utils import read_config, setup_logging
 
 log = setup_logging()
-
 def main():
     log.info("Preparing data for analysis ...")
     cfg = read_config('config/prepare_data_cfg.yaml')
@@ -21,6 +20,9 @@ def main():
     # Check and handle blank values in vital fields
     vital_fields = ['transparency_report_fkey', 'entity_map_fkey', 'auditor_fkey', 'trans_report_auditor_state']
     transparency_data = check_and_handle_blanks(transparency_data, vital_fields)
+
+    # Ensure rows with NUMBER_OF_DISCLOSED_PIES > 0
+    transparency_data = filter_disclosed_pies(transparency_data)
 
     # Check for duplicates in entity_map_fkey
     check_for_joint_audits(transparency_data)
@@ -55,6 +57,21 @@ def check_and_handle_blanks(df, fields):
     final_count = len(df)
     log.info(f"Rows removed due to blank values: {initial_count - final_count}")
     return df
+
+
+def filter_disclosed_pies(df):
+    """
+    Ensure that only rows with NUMBER_OF_DISCLOSED_PIES > 0 are included.
+    """
+    initial_count = len(df)
+    df = df[df['number_of_disclosed_pies'] > 0]
+    filtered_count = initial_count - len(df)
+    if filtered_count > 0:
+        log.info(f"{filtered_count} rows were filtered out because NUMBER_OF_DISCLOSED_PIES <= 0.")
+    else:
+        log.info("No rows were filtered out based on NUMBER_OF_DISCLOSED_PIES.")
+    return df
+
 
 def check_for_joint_audits(df):
     """

### 2024-11-16 code/python/prepare_data.py
prepare data final

@@ -17,6 +17,12 @@ def main():
     initial_obs_count = len(transparency_data)
     log.info(f"Initial number of observations after pulling data: {initial_obs_count}")
 
+    # Standardize Greece abbreviation
+    transparency_data = standardize_greece_abbreviation(transparency_data)
+
+    # Verify disclosed PIEs alignment with listed entities
+    transparency_data = verify_disclosed_pies_alignment(transparency_data)
+
     # Check and handle blank values in vital fields
     vital_fields = ['transparency_report_fkey', 'entity_map_fkey', 'auditor_fkey', 'trans_report_auditor_state']
     transparency_data = check_and_handle_blanks(transparency_data, vital_fields)
@@ -43,6 +49,49 @@ def main():
     log.info(f"Prepared data saved to {cfg['prepared_data_save_path']}")
 
     log.info("Preparing data for analysis ... Done!")
+
+def standardize_greece_abbreviation(df):
+    """
+    Replace 'GR' with 'EL' for Greece in the trans_report_auditor_state column.
+    """
+    greece_count = df[df['trans_report_auditor_state'] == 'GR'].shape[0]
+    if greece_count > 0:
+        log.info(f"Found {greece_count} rows with 'GR' abbreviation for Greece. Replacing with 'EL'.")
+        df['trans_report_auditor_state'] = df['trans_report_auditor_state'].replace('GR', 'EL')
+    else:
+        log.info("No 'GR' abbreviation found for Greece. No replacements made.")
+    return df
+
+def verify_disclosed_pies_alignment(df):
+    """
+    Verify that the number of unique entity_map_fkey matches number_of_disclosed_pies
+    for each transparency_report_fkey.
+    """
+    # Group by transparency_report_fkey and count unique entity_map_fkey
+    entity_counts = df.groupby('transparency_report_fkey')['entity_map_fkey'].nunique().reset_index()
+    entity_counts.columns = ['transparency_report_fkey', 'entity_count']
+
+    # Merge with original dataframe to compare counts
+    merged_df = df.merge(entity_counts, on='transparency_report_fkey', how='left')
+
+    # Check for mismatches
+    overreported = merged_df[merged_df['entity_count'] > merged_df['number_of_disclosed_pies']]
+    underreported = merged_df[merged_df['entity_count'] < merged_df['number_of_disclosed_pies']]
+
+    if not overreported.empty:
+        log.warning(f"{len(overreported)} transparency reports include more entities than disclosed PIEs.")
[…]

### 2024-11-17 code/python/do_analysis old.py
do analysis

@@ -1,0 +1,266 @@
+# We start by loading the libraries that we will use in this analysis.
+import pickle
+import pandas as pd
+from utils import read_config, setup_logging
+
+# Set up logging
+log = setup_logging()
+
+def main():
+    log.info("Performing main analysis...")
+    
+    # Load the configuration file
+    cfg = read_config('config/do_analysis_cfg.yaml')
+    
+    # Load the prepared financial data using the path from the config file
+    financial_data = load_data(cfg['prepared_data_save_path'])
+
+    # Calculate EM1
+    country_em1, summary_stats_em1 = calculate_em1(financial_data)
+
+    # Calculate EM2
+    country_em2, summary_stats_em2 = calculate_em2(financial_data)
+
+    # Calculate EM3
+    country_em3, summary_stats_em3 = calculate_em3(financial_data)
+    
+    # Calculate EM4
+    country_em4, summary_stats_em4 = calculate_em4(financial_data)
+
+    # Calculate Aggregate Earnings Management Score and create final table
+    final_table = calculate_aggregate(country_em1, country_em2, country_em3, country_em4)
+
+    # Create the final combined table with both metrics and summary statistics
+    final_combined_table = create_final_combined_table(final_table)
+
+    # Save the final combined table using the path from the config file
+    save_results(final_combined_table, cfg['results'])
+
+    log.info("Performing main analysis...Done!")
+
+def load_data(data_path):
+    """
+    Load the prepared financial data from the specified path.
+    """
+    df = pd.read_csv(data_path)
+    return df
+
+def calculate_em1(df):
+    """
+    Calculate EM1 for each firm and then take the country-level median.
+    This function combines all necessary steps into one.
+    """
+    # Step 1: Calculate Accruals
+    # Group by unique firm (item6105) and calculate diffs within each firm
+    # If a firm does not report information on taxes payable or short-term debt,
+    # then the change in both variables is assumed to be zero (per paper).
+    df['delta_CA'] = df.groupby('item6105')['item2201'].diff()  # Change in total current assets
+    df['delta_Cash'] = df.groupby('item6105')['item2003'].diff()  # Change in cash and cash equivalents
[…]

### 2024-11-17 code/python/do_analysis.py

@@ -12,255 +12,120 @@ def main():
     # Load the configuration file
     cfg = read_config('config/do_analysis_cfg.yaml')
     
-    # Load the prepared financial data using the path from the config file
-    financial_data = load_data(cfg['prepared_data_save_path'])
-
-    # Calculate EM1
-    country_em1, summary_stats_em1 = calculate_em1(financial_data)
-
-    # Calculate EM2
-    country_em2, summary_stats_em2 = calculate_em2(financial_data)
-
-    # Calculate EM3
-    country_em3, summary_stats_em3 = calculate_em3(financial_data)
+    # Load the prepared data
+    prepared_data = load_data(cfg['prepared_data_save_path'])
     
-    # Calculate EM4
-    country_em4, summary_stats_em4 = calculate_em4(financial_data)
-
-    # Calculate Aggregate Earnings Management Score and create final table
-    final_table = calculate_aggregate(country_em1, country_em2, country_em3, country_em4)
-
-    # Create the final combined table with both metrics and summary statistics
-    final_combined_table = create_final_combined_table(final_table)
-
-    # Save the final combined table using the path from the config file
-    save_results(final_combined_table, cfg['results'])
-
-    log.info("Performing main analysis...Done!")
+    # Calculate market shares for Big 4, 10KAP, and CR4
+    market_shares = calculate_market_shares(prepared_data)
+    
+    # Save the aggregated market shares to a CSV file
+    save_market_shares(market_shares, cfg['aggregated_data_save_path'])
+    
+    log.info("Performing main analysis for Figure 3 replication... Done!")
 
 def load_data(data_path):
     """
-    Load the prepared financial data from the specified path.
+    Load the prepared transparency data from the specified path.
     """
+    log.info(f"Loading prepared data from {data_path}...")
     df = pd.read_csv(data_path)
     return df
 
-def calculate_em1(df):
+def calculate_market_shares(df):
     """
-    Calculate EM1 for each firm and then take the country-level median.
-    This function combines all necessary steps into one.
+    Calculate market shares for Big 4, 10KAP, and CR4 for each country and for the EU as a whole.
     """
-    # Step 1: Calculate Accruals
[…]

### 2024-11-17 code/python/do_analysis.py
do more analysis

@@ -12,7 +12,7 @@ def main():
     # Load the configuration file
     cfg = read_config('config/do_analysis_cfg.yaml')
     
-    # Load the prepared data
+    # Load the prepared data using the path from the config file
     prepared_data = load_data(cfg['prepared_data_save_path'])
     
     # Calculate market shares for Big 4, 10KAP, and CR4
@@ -21,7 +21,7 @@ def main():
     # Save the aggregated market shares to a CSV file
     save_market_shares(market_shares, cfg['aggregated_data_save_path'])
     
-    log.info("Performing main analysis for Figure 3 replication... Done!")
+    log.info("Performing main analysis...Done!")
 
 def load_data(data_path):
     """

### 2024-11-17 code/python/do_analysis.py
analysis

@@ -1,5 +1,5 @@
 # We start by loading the libraries that we will use in this analysis.
-import pickle
+import os
 import pandas as pd
 from utils import read_config, setup_logging
 
@@ -15,8 +15,14 @@ def main():
     # Load the prepared data using the path from the config file
     prepared_data = load_data(cfg['prepared_data_save_path'])
     
-    # Calculate market shares for Big 4, 10KAP, and CR4
-    market_shares = calculate_market_shares(prepared_data)
+    # Calculate market shares for each group
+    big4_shares = calculate_big4_market_share(prepared_data)
+    kap10_shares = calculate_kap10_market_share(prepared_data)
+    cr4_shares = calculate_cr4_market_share(prepared_data)
+    eu_shares = calculate_eu_level_market_shares(prepared_data)
+    
+    # Combine results into one DataFrame
+    market_shares = combine_market_shares(big4_shares, kap10_shares, cr4_shares, eu_shares)
     
     # Save the aggregated market shares to a CSV file
     save_market_shares(market_shares, cfg['aggregated_data_save_path'])
@@ -28,75 +34,60 @@ def load_data(data_path):
     Load the prepared transparency data from the specified path.
     """
     log.info(f"Loading prepared data from {data_path}...")
-    df = pd.read_csv(data_path)
-    return df
+    return pd.read_csv(data_path)
 
-def calculate_market_shares(df):
+def calculate_big4_market_share(df):
     """
-    Calculate market shares for Big 4, 10KAP, and CR4 for each country and for the EU as a whole.
+    Calculate market share for Big 4 by country.
     """
-    # Group by country and calculate total PIE audits
-    country_totals = (
-        df.groupby('trans_report_auditor_state')['number_of_disclosed_pies']
-        .sum()
-        .reset_index(name='total_pie_audits')
-    )
+    country_totals = df.groupby('trans_report_auditor_state')['number_of_disclosed_pies'].sum().reset_index()
+    country_totals.columns = ['trans_report_auditor_state', 'total_pie_audits']
     
-    # Big 4 Market Share
-    big4_shares = calculate_group_market_share(df, 'Big 4', country_totals)
+    big4_df = df[df['network_group'] == 'Big 4']
+    big4_totals = big4_df.groupby('trans_report_auditor_state')['number_of_disclosed_pies'].sum().reset_index()
[…]

### 2024-11-17 code/python/do_analysis.py
analysis

@@ -12,7 +12,7 @@ def main():
     # Load the configuration file
     cfg = read_config('config/do_analysis_cfg.yaml')
     
-    # Load the prepared data using the path from the config file
+    # Load the prepared transparency data using the path from the config file
     prepared_data = load_data(cfg['prepared_data_save_path'])
     
     # Calculate market shares for each group
@@ -33,8 +33,8 @@ def load_data(data_path):
     """
     Load the prepared transparency data from the specified path.
     """
-    log.info(f"Loading prepared data from {data_path}...")
-    return pd.read_csv(data_path)
+    df = pd.read_csv(data_path)
+    return df
 
 def calculate_big4_market_share(df):
     """

### 2024-11-17 code/python/do_analysis.py
CR4

@@ -40,54 +40,90 @@ def calculate_big4_market_share(df):
     """
     Calculate market share for Big 4 by country.
     """
-    country_totals = df.groupby('trans_report_auditor_state')['number_of_disclosed_pies'].sum().reset_index()
-    country_totals.columns = ['trans_report_auditor_state', 'total_pie_audits']
+    # Step 1: Calculate total PIE audits by counting rows per country
+    country_totals = df.groupby('trans_report_auditor_state').size().reset_index(name='total_pie_audits')
+    print("Country Totals (All Audits):")
+    print(country_totals.head(9))
     
+    # Step 2: Filter data for Big 4
     big4_df = df[df['network_group'] == 'Big 4']
-    big4_totals = big4_df.groupby('trans_report_auditor_state')['number_of_disclosed_pies'].sum().reset_index()
-    big4_totals.columns = ['trans_report_auditor_state', 'big4_pie_audits']
+    print("Filtered Big 4 Data:")
+    print(big4_df.head())
     
+    # Step 3: Calculate Big 4 PIE audits by counting rows per country
+    big4_totals = big4_df.groupby('trans_report_auditor_state').size().reset_index(name='big4_pie_audits')
+    print("Big 4 Totals (Audits by Country):")
+    print(big4_totals.head(10))
+    
+    # Step 4: Merge with country totals and calculate market share
     big4_shares = big4_totals.merge(country_totals, on='trans_report_auditor_state')
     big4_shares['big4_market_share'] = (big4_shares['big4_pie_audits'] / big4_shares['total_pie_audits']) * 100
+    print("Big 4 Market Shares:")
+    print(big4_shares.head(10))
     
+    # Step 5: Return final DataFrame
     return big4_shares[['trans_report_auditor_state', 'big4_market_share']]
 
 def calculate_kap10_market_share(df):
     """
-    Calculate market share for 10KAP by country.
+    Calculate market share for 10KAP by country, including Big 4 auditors.
     """
-    country_totals = df.groupby('trans_report_auditor_state')['number_of_disclosed_pies'].sum().reset_index()
-    country_totals.columns = ['trans_report_auditor_state', 'total_pie_audits']
+    # Step 1: Calculate total PIE audits by counting rows per country
+    country_totals = df.groupby('trans_report_auditor_state').size().reset_index(name='total_pie_audits')
+    print("Country Totals (All Audits):")
[…]

### 2024-11-17 code/python/do_analysis.py
overlap check

@@ -18,7 +18,7 @@ def main():
     # Calculate market shares for each group
     big4_shares = calculate_big4_market_share(prepared_data)
     kap10_shares = calculate_kap10_market_share(prepared_data)
-    cr4_shares = calculate_cr4_market_share(prepared_data)
+    cr4_shares = calculate_cr4_market_share(prepared_data, big4_shares)  # Pass big4_shares as an argument
     eu_shares = calculate_eu_level_market_shares(prepared_data)
     
     # Combine results into one DataFrame
@@ -92,9 +92,10 @@ def calculate_kap10_market_share(df):
     # Step 5: Return final DataFrame
     return kap10_shares[['trans_report_auditor_state', 'kap10_market_share']]
 
-def calculate_cr4_market_share(df):
+def calculate_cr4_market_share(df, big4_shares):
     """
-    Calculate the CR4 market share for the four largest audit firms in each country.
+    Calculate the CR4 market share for the four largest audit firms in each country,
+    and additionally check for overlap with Big 4 market share.
     """
     # Step 1: Count statutory audits by firm within each country
     firm_totals = (
@@ -123,8 +124,15 @@ def calculate_cr4_market_share(df):
     print("CR4 Market Shares (By Country):")
     print(cr4_shares.head(9))
 
-    # Step 5: Return final DataFrame
-    return cr4_shares[['trans_report_auditor_state', 'cr4_market_share']]
+    # Step 5: Merge with Big 4 market shares to check overlap
+    cr4_shares = cr4_shares.merge(big4_shares, on='trans_report_auditor_state', how='left')
+    cr4_shares['overlap_with_big4'] = cr4_shares['cr4_market_share'] == cr4_shares['big4_market_share']
+
+    print("CR4 Market Shares with Overlap Check (By Country):")
+    print(cr4_shares[['trans_report_auditor_state', 'cr4_market_share', 'big4_market_share', 'overlap_with_big4']])
+
+    # Step 6: Return final DataFrame with overlap flag
+    return cr4_shares[['trans_report_auditor_state', 'cr4_market_share', 'overlap_with_big4']]
 
 def calculate_eu_level_market_shares(df):
     """

### 2024-11-17 code/python/do_analysis.py
finalized analysis

@@ -95,7 +95,7 @@ def calculate_kap10_market_share(df):
 def calculate_cr4_market_share(df, big4_shares):
     """
     Calculate the CR4 market share for the four largest audit firms in each country,
-    and additionally check for overlap with Big 4 market share.
+    and check for overlap with Big 4 market share.
     """
     # Step 1: Count statutory audits by firm within each country
     firm_totals = (
@@ -131,25 +131,56 @@ def calculate_cr4_market_share(df, big4_shares):
     print("CR4 Market Shares with Overlap Check (By Country):")
     print(cr4_shares[['trans_report_auditor_state', 'cr4_market_share', 'big4_market_share', 'overlap_with_big4']])
 
-    # Step 6: Return final DataFrame with overlap flag
+    # Step 6: Calculate the number of countries with overlap
+    overlap_count = cr4_shares['overlap_with_big4'].sum()
+    print(f"Number of countries with CR4 overlapping with Big 4: {overlap_count}")
+
+    # Step 7: Return final DataFrame with overlap flag
     return cr4_shares[['trans_report_auditor_state', 'cr4_market_share', 'overlap_with_big4']]
 
 def calculate_eu_level_market_shares(df):
     """
-    Calculate EU-level market shares for Big 4, 10KAP, and CR4.
+    Calculate EU-level market shares for Big 4, 10KAP, and CR4 using statutory audit counts.
     """
-    eu_totals = df['number_of_disclosed_pies'].sum()
-    big4_total = df[df['network_group'] == 'Big 4']['number_of_disclosed_pies'].sum()
-    kap10_total = df[df['network_group'] == '10KAP']['number_of_disclosed_pies'].sum()
-    
-    firm_totals = df.groupby('auditor_fkey')['number_of_disclosed_pies'].sum().nlargest(4).sum()
-    
+    # Step 1: Calculate total statutory audits in the EU by counting rows
+    eu_totals = len(df)
+    print(f"Total Statutory Audits in the EU: {eu_totals}")
+
+    # Step 2: Filter data for Big 4 and calculate total audits
+    big4_df = df[df['network_group'] == 'Big 4']
+    big4_total = len(big4_df)
+    print(f"Total Big 4 Audits in the EU: {big4_total}")
+
+    # Step 3: Filter data for 10KAP (including Big 4) and calculate total audits
[…]

### 2024-11-17 code/python/do_analysis.py
plot final

@@ -1,6 +1,7 @@
 # We start by loading the libraries that we will use in this analysis.
 import os
 import pandas as pd
+import matplotlib.pyplot as plt
 from utils import read_config, setup_logging
 
 # Set up logging
@@ -26,6 +27,9 @@ def main():
     
     # Save the aggregated market shares to a CSV file
     save_market_shares(market_shares, cfg['aggregated_data_save_path'])
+
+    # Plot the results
+    plot_market_shares(market_shares, cfg['figure_save_path'])
     
     log.info("Performing main analysis...Done!")
 
@@ -199,6 +203,38 @@ def save_market_shares(df, save_path):
     os.makedirs(os.path.dirname(save_path), exist_ok=True)
     df.to_csv(save_path, index=False)
     log.info(f"Market shares saved to {save_path}.")
+
+def plot_market_shares(df, save_path):
+    """
+    Plot the market shares for Big 4, 10KAP, and CR4.
+    """
+    # Filter out the EU-level row for country-specific plotting
+    country_data = df[df['trans_report_auditor_state'] != 'EU']
+    
+    # Set up the bar chart
+    fig, ax = plt.subplots(figsize=(14, 7))
+    x = country_data['trans_report_auditor_state']
+    width = 0.3
+    
+    # Plot Big 4, CR4, and 10KAP market shares
+    ax.bar(x, country_data['big4_market_share'], width, label='Big 4', align='center')
+    ax.bar(x, country_data['cr4_market_share'], width, label='CR4', align='edge')
+    ax.bar(x, country_data['kap10_market_share'], width, label='10KAP', align='edge', color='lightgrey', alpha=0.7)
+    
+    # Add labels and title
+    ax.set_xlabel('Country')
+    ax.set_ylabel('Market Share (%)')
+    ax.set_title('Audit Firms\' Market Share in Number of PIE Statutory Audits (2021)')
+    ax.legend()
+    ax.set_xticks(range(len(x)))
+    ax.set_xticklabels(x, rotation=90)
+    
+    # Save the figure
+    os.makedirs(os.path.dirname(save_path), exist_ok=True)
+    plt.tight_layout()
+    plt.savefig(save_path)
+    log.info(f"Figure saved to {save_path}.")
+    
 
 if __name__ == "__main__":
     main()

### 2024-11-17 code/python/do_analysis.py
plot

@@ -31,7 +31,7 @@ def main():
     # Plot the results
     plot_market_shares(market_shares, cfg['figure_save_path'])
     
-    log.info("Performing main analysis...Done!")
+    log.info("Performing main analysis and plotting...Done!")
 
 def load_data(data_path):
     """
@@ -204,37 +204,43 @@ def save_market_shares(df, save_path):
     df.to_csv(save_path, index=False)
     log.info(f"Market shares saved to {save_path}.")
 
-def plot_market_shares(df, save_path):
+def plot_market_shares(market_shares, save_path):
     """
-    Plot the market shares for Big 4, 10KAP, and CR4.
+    Create a bar chart for Big 4, CR4, and 10KAP market shares by country (including EU).
     """
-    # Filter out the EU-level row for country-specific plotting
-    country_data = df[df['trans_report_auditor_state'] != 'EU']
+    # Sort by country, placing EU at the end for clarity
+    market_shares = market_shares.sort_values(by='trans_report_auditor_state', key=lambda col: col != 'EU')
+
+    # Extract data for plotting
+    countries = market_shares['trans_report_auditor_state']
+    big4 = market_shares['big4_market_share']
+    cr4 = market_shares['cr4_market_share']
+    kap10 = market_shares['kap10_market_share']
     
-    # Set up the bar chart
+    # Plot settings
+    x = range(len(countries))
+    bar_width = 0.25
+
     fig, ax = plt.subplots(figsize=(14, 7))
-    x = country_data['trans_report_auditor_state']
-    width = 0.3
+    ax.bar(x, big4, width=bar_width, label='Big 4', color='blue', align='center')
+    ax.bar([i + bar_width for i in x], cr4, width=bar_width, label='CR4', color='orange', align='center')
+    ax.bar([i + 2 * bar_width for i in x], kap10, width=bar_width, label='10KAP', color='grey', align='center')
     
-    # Plot Big 4, CR4, and 10KAP market shares
-    ax.bar(x, country_data['big4_market_share'], width, label='Big 4', align='center')
-    ax.bar(x, country_data['cr4_market_share'], width, label='CR4', align='edge')
-    ax.bar(x, country_data['kap10_market_share'], width, label='10KAP', align='edge', color='lightgrey', alpha=0.7)
-    
-    # Add labels and title
-    ax.set_xlabel('Country')
[…]

### 2024-11-17 code/python/do_analysis.py
adjust plot

@@ -206,27 +206,40 @@ def save_market_shares(df, save_path):
 
 def plot_market_shares(market_shares, save_path):
     """
-    Create a bar chart for Big 4, CR4, and 10KAP market shares by country (including EU).
+    Create a patterned bar chart for Big 4, CR4, and 10KAP market shares by country (including EU).
     """
-    # Sort by country, placing EU at the end for clarity
-    market_shares = market_shares.sort_values(by='trans_report_auditor_state', key=lambda col: col != 'EU')
-
+    # Sort the market shares to match the desired country order
+    countries_order = ['DK', 'CY', 'FI', 'SE', 'LU', 'BE', 'EE', 'AT', 'NO', 'NL', 'IT', 
+                       'ES', 'IE', 'DE', 'MT', 'LT', 'HU', 'EU', 'LV', 'CZ', 'HR', 'SI', 
+                       'SK', 'FR', 'PL', 'PT', 'EL', 'RO', 'BG']
+    market_shares['sort_order'] = market_shares['trans_report_auditor_state'].apply(lambda x: countries_order.index(x))
+    market_shares = market_shares.sort_values(by='sort_order')
+    
     # Extract data for plotting
     countries = market_shares['trans_report_auditor_state']
     big4 = market_shares['big4_market_share']
     cr4 = market_shares['cr4_market_share']
     kap10 = market_shares['kap10_market_share']
     
-    # Plot settings
+    # Define bar width and x-coordinates
     x = range(len(countries))
     bar_width = 0.25
 
+    # Plot settings
     fig, ax = plt.subplots(figsize=(14, 7))
-    ax.bar(x, big4, width=bar_width, label='Big 4', color='blue', align='center')
-    ax.bar([i + bar_width for i in x], cr4, width=bar_width, label='CR4', color='orange', align='center')
-    ax.bar([i + 2 * bar_width for i in x], kap10, width=bar_width, label='10KAP', color='grey', align='center')
-    
-    # Format the plot
+
+    # Add bars with appropriate colors and patterns
+    ax.bar(x, big4, width=bar_width, label='Big 4', color='darkblue', align='center')
+    ax.bar([i + bar_width for i in x], cr4, width=bar_width, label='CR4', color='lightblue', align='center')
+    ax.bar([i + 2 * bar_width for i in x], kap10, width=bar_width, label='10KAP', color='gray', align='center')
+
+    # Highlight EU in red for all groups
[…]

### 2024-11-17 code/python/do_analysis.py
graphics update

@@ -228,24 +228,21 @@ def plot_market_shares(market_shares, save_path):
     # Plot settings
     fig, ax = plt.subplots(figsize=(14, 7))
 
-    # Add bars with appropriate colors and patterns
+    # Add bars with appropriate colors
     ax.bar(x, big4, width=bar_width, label='Big 4', color='darkblue', align='center')
     ax.bar([i + bar_width for i in x], cr4, width=bar_width, label='CR4', color='lightblue', align='center')
     ax.bar([i + 2 * bar_width for i in x], kap10, width=bar_width, label='10KAP', color='gray', align='center')
 
-    # Highlight EU in red for all groups
-    eu_index = countries[countries == 'EU'].index[0]
-    ax.bar(eu_index, big4.iloc[eu_index], width=bar_width, color='red', align='center')
-    ax.bar(eu_index + bar_width, cr4.iloc[eu_index], width=bar_width, color='red', align='center')
-    ax.bar(eu_index + 2 * bar_width, kap10.iloc[eu_index], width=bar_width, color='red', align='center')
-
     # Add labels, title, and legend
     ax.set_xlabel('Country', fontsize=12)
     ax.set_ylabel('Market Share (%)', fontsize=12)
-    ax.set_title("Audit Firms' Market Share in Number of PIE Statutory Audits (2021)", fontsize=14)
+    ax.set_title("Replication of Figure 3: Audit Firms' Market Share in Number of PIE Statutory Audits (2021)", fontsize=14)
     ax.set_xticks([i + bar_width for i in x])
     ax.set_xticklabels(countries, rotation=90)
     ax.legend()
+
+    # Position the legend below the graph
+    ax.legend(loc='upper center', bbox_to_anchor=(0.5, -0.15), fancybox=True, shadow=True, ncol=3)
     
     # Save the figure
     os.makedirs(os.path.dirname(save_path), exist_ok=True)

### 2024-11-17 code/python/do_analysis.py
EU graphic match

@@ -229,9 +229,17 @@ def plot_market_shares(market_shares, save_path):
     fig, ax = plt.subplots(figsize=(14, 7))
 
     # Add bars with appropriate colors
-    ax.bar(x, big4, width=bar_width, label='Big 4', color='darkblue', align='center')
-    ax.bar([i + bar_width for i in x], cr4, width=bar_width, label='CR4', color='lightblue', align='center')
-    ax.bar([i + 2 * bar_width for i in x], kap10, width=bar_width, label='10KAP', color='gray', align='center')
+    for i, country in enumerate(countries):
+        if country == 'EU':
+            # Highlight EU with red color
+            ax.bar(i, big4.iloc[i], width=bar_width, color='darkred', align='center')
+            ax.bar(i + bar_width, cr4.iloc[i], width=bar_width, color='red', align='center')
+            ax.bar(i + 2 * bar_width, kap10.iloc[i], width=bar_width, color='lightcoral', align='center')
+        else:
+            # Default colors for other countries
+            ax.bar(i, big4.iloc[i], width=bar_width, color='darkblue', align='center')
+            ax.bar(i + bar_width, cr4.iloc[i], width=bar_width, color='lightblue', align='center')
+            ax.bar(i + 2 * bar_width, kap10.iloc[i], width=bar_width, color='gray', align='center')
 
     # Add labels, title, and legend
     ax.set_xlabel('Country', fontsize=12)

### 2024-11-17 code/python/do_analysis.py
legend update

@@ -231,15 +231,15 @@ def plot_market_shares(market_shares, save_path):
     # Add bars with appropriate colors
     for i, country in enumerate(countries):
         if country == 'EU':
-            # Highlight EU with red color
-            ax.bar(i, big4.iloc[i], width=bar_width, color='darkred', align='center')
-            ax.bar(i + bar_width, cr4.iloc[i], width=bar_width, color='red', align='center')
-            ax.bar(i + 2 * bar_width, kap10.iloc[i], width=bar_width, color='lightcoral', align='center')
+            # Highlight EU with different red shades
+            ax.bar(i, big4.iloc[i], width=bar_width, color='darkred', align='center', label='Big 4 (EU)' if i == 0 else "")
+            ax.bar(i + bar_width, cr4.iloc[i], width=bar_width, color='red', align='center', label='CR4 (EU)' if i == 0 else "")
+            ax.bar(i + 2 * bar_width, kap10.iloc[i], width=bar_width, color='lightcoral', align='center', label='10KAP (EU)' if i == 0 else "")
         else:
             # Default colors for other countries
-            ax.bar(i, big4.iloc[i], width=bar_width, color='darkblue', align='center')
-            ax.bar(i + bar_width, cr4.iloc[i], width=bar_width, color='lightblue', align='center')
-            ax.bar(i + 2 * bar_width, kap10.iloc[i], width=bar_width, color='gray', align='center')
+            ax.bar(i, big4.iloc[i], width=bar_width, color='darkblue', align='center', label='Big 4' if i == 0 else "")
+            ax.bar(i + bar_width, cr4.iloc[i], width=bar_width, color='lightblue', align='center', label='CR4' if i == 0 else "")
+            ax.bar(i + 2 * bar_width, kap10.iloc[i], width=bar_width, color='gray', align='center', label='10KAP' if i == 0 else "")
 
     # Add labels, title, and legend
     ax.set_xlabel('Country', fontsize=12)
@@ -247,9 +247,8 @@ def plot_market_shares(market_shares, save_path):
     ax.set_title("Replication of Figure 3: Audit Firms' Market Share in Number of PIE Statutory Audits (2021)", fontsize=14)
     ax.set_xticks([i + bar_width for i in x])
     ax.set_xticklabels(countries, rotation=90)
-    ax.legend()
-
-    # Position the legend below the graph
+    
+    # Define legend below the graph
[…]

### 2024-11-17 code/python/do_analysis.py
finalized analysis

@@ -244,7 +244,7 @@ def plot_market_shares(market_shares, save_path):
     # Add labels, title, and legend
     ax.set_xlabel('Country', fontsize=12)
     ax.set_ylabel('Market Share (%)', fontsize=12)
-    ax.set_title("Replication of Figure 3: Audit Firms' Market Share in Number of PIE Statutory Audits (2021)", fontsize=14)
+    ax.set_title("Replication of Figure 3 from EC Report (2024): Audit Firms' Market Share in Number of PIE Statutory Audits (2021)", fontsize=14)
     ax.set_xticks([i + bar_width for i in x])
     ax.set_xticklabels(countries, rotation=90)
     

### 2024-11-17 doc/presentation.qmd
presentation update

@@
----
+--- ## Not rendered in this project; to be used as draft
 title: |

### 2024-11-17 doc/paper.qmd
Abstract update

@@
 title: |
-  | Earnings Management and Investor Protection:
-  | Accounting Reading Group - Assignment III\vspace{1cm}
+  | Audit market concentration in the EU:
+  | Corporate Decision-Making and Quantitative Analysis - Assignment I\vspace{1cm}
 author:
@@
   | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide an infrastructure for open science-oriented empirical projects.
-Leveraging Worldscope data on financial information, this repository showcases a reproducible workflow that integrates Python scripts for data analysis.
-The project’s output demonstrates a comprehensive application of skills to replicate and extend the findings from the seminal paper by Leuz, Nanda, and Wysocki (2003), particularly in providing descriptive statistics for four individual earnings management measures, as well as an aggregate earnings management score across various countries.
+Leveraging Audit Analytics database on Transparency Report data, this repository showcases a reproducible workflow that integrates Python scripts for data preparation, analysis, and visualization.
+The project replicates and extends findings from the European Commission’s 2024 report by analyzing and visualizing audit firms’ Big 4, CR4, and 10KAP  market shares in the number of PIE statutory audits for 2021 across EU countries and at the EU level.
+The results include a detailed breakdown of market concentration trends, offering insights into the dominance of major audit firms across different jurisdictions.
 In doing so, it documents and discusses the research design choices made and the variations between the original and reproduced results.

### 2024-11-17 doc/paper.qmd
paper establishment

@@
   | This project uses the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) to provide an infrastructure for open science-oriented empirical projects.
-Leveraging Audit Analytics database on Transparency Report data, this repository showcases a reproducible workflow that integrates Python scripts for data preparation, analysis, and visualization.
-The project replicates and extends findings from the European Commission’s 2024 report by analyzing and visualizing audit firms’ Big 4, CR4, and 10KAP  market shares in the number of PIE statutory audits for 2021 across EU countries and at the EU level.
+Leveraging the Audit Analytics database on Transparency Report data, this repository showcases a reproducible workflow that integrates Python scripts for data preparation, analysis, and visualization.
+The project replicates and extends findings from the European Commission’s 2024 report by analyzing and visualizing Big 4, CR4, and 10KAP audit firms’ market shares in the number of public interest entity statutory audits for 2021 across EU countries and at the EU level.
 The results include a detailed breakdown of market concentration trends, offering insights into the dominance of major audit firms across different jurisdictions.
-In doing so, it documents and discusses the research design choices made and the variations between the original and reproduced results.
-This code base, adapted from TREAT, provides an overview of how the template can be used for this specific project and how to structure a reproducible empirical research project.
+In doing so, the project documents and discusses the research design choices made and the variations between the original and reproduced results.
+This code base, adapted from TREAT, provides an overview of how the template can be used for this specific project and demonstrates how to structure a reproducible empirical research project.
   | \vspace{6cm}
@@
 \setlength{\parindent}{0em}
-
-**CFO**: Operating Cash Flow  
-**EM**: Earnings Management  
+**EC**: European Commission  
+**EU**: European Union  
 **IDE**: Integrated Development Environment  
-**SIC**: Standard Industrial Classification  
+**NCA**:: National Competent Authority  
+**PIE**: Public interest entity  
+**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
 **UK**: United Kingdom   
-**US**: United States  
 **WRDS**: Wharton Research Data Services
@@
 
-The original study by @Leuz_2003 examines systematic differences in earnings management (EM) across 31 countries and proposes that these differences are influenced by the level of investor protection.
-Specifically, the paper argues that stronger investor protection reduces earnings management by limiting insiders’ ability to acquire private control benefits.
+The original report by @EC_Report_2024 examines trends in the European Union (EU) market for statutory audits of public interest entities (PIEs) from 2019 to 2021, with a focus on market concentration, audit quality risks, and the effectiveness of audit committees.
+This report is the third check-in on how the European Union's (EU’s) auditing standards and market structures are evolving since the *Article 27 of Regulation (EU) No 537/2014* [-@EU_Regulation_537_2014] was introduced.
+A statutory audit is a legally mandated evaluation of financial records aimed at certifying the accuracy of companies’ financial statements, thereby enhancing stakeholder confidence and supporting the orderly functioning of markets [@EC_Auditing_Website].
+The report by @EC_Report_2024 argues that while statutory audits play a crucial role in fostering confidence in financial markets, persistent market concentration and quality deficiencies highlight the need for strengthened oversight and improved transparency to ensure the resilience and reliability of audit practices across the EU.
 
-This paper presents the replication of Table 2 Panel A from the seminal study by @Leuz_2003.
-The original table provides descriptive statistics for four individual earnings management measures and the aggregate earnings management score across various countries, highlighting significant differences in EM practices between Continental Europe, Asia, and Anglo-American countries.
+This paper replicates Figure 3 from the European Commission’s [-@EC_Report_2024] report, which visualizes the market shares of audit firms — Big 4, CR4, and 10KAP — by the number of PIE statutory audits conducted in 2021 across EU countries and at the EU level.
+The replication provides insights into market concentration and the dominance of major audit firms, reflecting variations in audit practices and competition across Europe.
 
-The project uses data from the Worldscope Database via Wharton Research Data Services (WRDS), focusing on fiscal years 1990 to 1999 and covering multiple countries.
-The replication aims to follow the research design by @Leuz_2003 as closely as possible with the available data.
-It employs four individual earnings management measures: earnings smoothing using accruals (EM1), the correlation between changes in accruals and cash flows (EM2), the magnitude of accruals (EM3), and the ratio of small profits to small losses (EM4).
-These measures are combined to create an aggregate earnings management score, providing a comprehensive overview of EM practices.
+The project uses data from the Audit Analytics database on Transparency Reports via Wharton Research Data Services (WRDS) for the year 2021, covering PIEs across EU countries.
+The replication closely follows the methodology outlined by @EC_Report_2024.
+It focuses on three key market concentration measures: the market share of the Big 4 audit firms, the CR4 market share for the largest four firms in each country, and the 10KAP market share, which includes the Big 4 and additional major networks.
+These measures provide a detailed assessment of audit firm dominance and competition, offering valuable insights into the structure and dynamics of the statutory audit market in Europe.
 
-The analysis includes pulling and filtering the relevant financial data, identifying relevant companies and countries based on the authors’ requirements, calculating the EM measures for each country, and examining the relationship between each country’s aggregate earnings management score and its ranking.
-Explicit assumptions were made whenever the paper by @Leuz_2003 was unclear on how to proceed.
-The entire Python computation code for EM measures is available in `code/python/do_analysis.py` for detailed review if necessary.
-This paper focuses on the replication process and discussion of the results.
+The analysis involves gathering and filtering Transparency Report data, identifying relevant PIEs and EU countries, and calculating the Big 4, CR4, and 10KAP market shares for each country as well as at the EU level.
+Explicit assumptions were made whenever the document by @EC_Report_2024 was unclear on how to proceed.
+The entire Python computation code for market share calculations is available in `code/python/do_analysis.py` for detailed review if necessary.
+This paper focuses on the replication process, presenting visualizations, and discussing the results.
 
-The replicated table is shown below, followed by the research design choices and assumptions in @sec-research_design_assumptions, documentation of the replication steps in @sec-replication_steps, a detailed results comparison in @sec-results, and concluding remarks in @sec-conclusion.
+The replicated Figure 3 from original report by @EC_Report_2024 is shown below, followed by the research design choices and assumptions in @sec-research_design_assumptions, documentation of the replication steps in @sec-replication_steps, a detailed results comparison in @sec-results, and concluding remarks in @sec-conclusion.
 

### 2024-11-17 doc/paper.qmd
visualisation done

@@
 
-```{python}
-#| echo: false
-#| output: false
-
-import pickle
-import pandas as pd
-
-# Load the results from the existing pickle file for table_1
-with open('../data/generated/table_1.pickle', 'rb') as f:
-    results = pickle.load(f)
-
-# Load the results from the new pickle file for final_table and summary_stats
-with open('../output/em_results.pickle', 'rb') as f:
-    em_results = pickle.load(f)
-
-def escape_special_characters(df):
-    df.columns = [col.replace('_', '\\_').replace('#', '\\#') for col in df.columns]
-    return df
-
-# Custom formatter function
-def custom_float_format(x):
-    return f"{x:.3f}".lstrip('0') if pd.notnull(x) else ""
-
-def prep_latex_table(df, caption=None, label=None, fontsize='\\fontsize{11}{13}\\selectfont', description=None):
-    df = escape_special_characters(df.reset_index(drop=True))  # Drop the index column from output tables for better readability
-    num_columns = len(df.columns)
-    column_format = 'l' + 'r' * (num_columns - 1)
-
-    latex_table = df.to_latex(
-        column_format=column_format,
-        index=False, 
-        float_format=custom_float_format,
-        escape=False  # Important to disable escaping so LaTeX commands in column names are respected
-    )
-
-    # Construct description to the table
-    full_caption = f"{caption}\\\\\n\\textnormal{{{description}}}" if description else caption
-
-    latex_table_lines = [
-        "\\begin{table}[htbp]",
-        "\\centering",
-        fontsize,  # Set font size for the table
-        latex_table,
-        f"\\caption{{{full_caption}}}" if full_caption else "",
-        f"\\label{{{label}}}" if label else "",
-        "\\end{table}"
-    ]
-    return "\n".join(line for line in latex_table_lines if line)
-```
-
 \pagebreak
@@
 
-The replicated Figure 3 from original report by @EC_Report_2024 is shown below, followed by the research design choices and assumptions in @sec-research_design_assumptions, documentation of the replication steps in @sec-replication_steps, a detailed results comparison in @sec-results, and concluding remarks in @sec-conclusion.
+The replicated figure from original report by @EC_Report_2024 is shown below, followed by the research design choices and assumptions in @sec-research_design_assumptions, documentation of the replication steps in @sec-replication_steps, a detailed results comparison in @sec-results, and concluding remarks in @sec-conclusion.
 
-Table 1 provides descriptive statistics for the four individual earnings management measures as well as the aggregate earnings management score, where countries are sorted in descending order based on their aggregate score.
+Figure 1 illustrates the market shares of audit firms — Big 4, CR4, and 10KAP — for the number of PIE statutory audits conducted in 2021.
+Countries are sorted in a specific order reflecting the replication of Figure 3 from the European Commission’s 2024 report.
 The discussion of replication findings is presented in @sec-results.
 ```{python}
-#| label: final_combined_table
+#| label: fig-market-shares
+#| fig-cap: "Replicated Figure 3: Audit Firms' Market Shares in the EU for 2021"
+#| fig-align: center
 #| echo: false
-#| output: asis
 
-# Rename the columns to ensure consistency with paper
-em_results['final_combined_table'].rename(columns={
-    'item6026': 'Country',
-    'EM1': 'EM1 (−)',
-    'EM2': 'EM2 (−)',
-    'EM3': 'EM3 (+)',
-    'EM4': 'EM4 (+)',
-    'Aggregate_EM_Score': 'Aggregate EM score'
-}, inplace=True)
+import pickle
+import matplotlib.pyplot as plt
 
-print(prep_latex_table(
-    em_results['final_combined_table'], 
-    caption='Replicated table - Country scores for earnings management measures (Sorted by aggregate earnings management)',
-    label='tab:final_combined_table',
-    description = (
-    "\\fontsize{10pt}{12pt}\\selectfont " 
-    "The variables are computed from 123,469 firm-year observations for fiscal years 1990 to 1999 across 31 countries and 18,040 non-financial firms.
-Data are obtained from the Worldscope Database (July 2024 version).
-The definitions and calculations of EM1-EM4, as well as the aggregate score, are thoroughly presented in Section 4.
-The sign in the column heading indicates whether higher scores for the respective EM measure imply more earnings management (+) or less earnings management (−)."
-    )
+# Load the figure from the pickle file
+with open('../output/figure3_market_shares.pickle', 'rb') as f:
+    fig = pickle.load(f)
 
-))
+# Display the figure
+fig
 ```
@@
 Despite the differences in absolute numbers, the relative trends across countries align with the original paper.
-
-```{python}
-#| label: table_1 
-#| echo: false
-#| output: asis
-
-print(prep_latex_table(
-    results['table_1'],
-    caption='Replicated table - Descriptive statistics: Number of firm-year observations per country',
-    label='tab:table_1'
-))
-```
 

### 2024-11-17 code/python/do_analysis.py

@@ -1,5 +1,6 @@
 # We start by loading the libraries that we will use in this analysis.
 import os
+import pickle
 import pandas as pd
 import matplotlib.pyplot as plt
 from utils import read_config, setup_logging
@@ -206,11 +207,12 @@ def save_market_shares(df, save_path):
 
 def plot_market_shares(market_shares, save_path):
     """
-    Create a patterned bar chart for Big 4, CR4, and 10KAP market shares by country (including EU).
+    Create a patterned bar chart for Big 4, CR4, and 10KAP market shares by country (including EU),
+    and save the figure as both PNG and pickle formats.
     """
     # Sort the market shares to match the desired country order
-    countries_order = ['DK', 'CY', 'FI', 'SE', 'LU', 'BE', 'EE', 'AT', 'NO', 'NL', 'IT', 
-                       'ES', 'IE', 'DE', 'MT', 'LT', 'HU', 'EU', 'LV', 'CZ', 'HR', 'SI', 
+    countries_order = ['DK', 'CY', 'FI', 'SE', 'LU', 'BE', 'EE', 'AT', 'NO', 'NL', 'IT',
+                       'ES', 'IE', 'DE', 'MT', 'LT', 'HU', 'EU', 'LV', 'CZ', 'HR', 'SI',
                        'SK', 'FR', 'PL', 'PT', 'EL', 'RO', 'BG']
     market_shares['sort_order'] = market_shares['trans_report_auditor_state'].apply(lambda x: countries_order.index(x))
     market_shares = market_shares.sort_values(by='sort_order')
@@ -251,13 +253,19 @@ def plot_market_shares(market_shares, save_path):
     # Define legend below the graph
     ax.legend(loc='upper center', bbox_to_anchor=(0.5, -0.15), fancybox=True, shadow=True, ncol=3)
     
-    # Save the figure
+    # Save the figure as a PNG
     os.makedirs(os.path.dirname(save_path), exist_ok=True)
     plt.tight_layout()
     plt.savefig(save_path)
     log.info(f"Figure saved to {save_path}.")
+    
+    # Save the figure as a pickle file
+    pickle_path = save_path.replace('.png', '.pickle')
+    with open(pickle_path, 'wb') as f:
+        pickle.dump(fig, f)
+    log.info(f"Figure saved as pickle to {pickle_path}.")
+    
     plt.show()
-
     
 if __name__ == "__main__":
     main()

### 2024-11-18 doc/paper.qmd
final commit

@@
 
-\pagebreak
-
 # List of Abbreviations
@@
 **IDE**: Integrated Development Environment  
-**NCA**:: National Competent Authority  
+**NCA**: National Competent Authority  
 **PIE**: Public interest entity  
@@
 The original report by @EC_Report_2024 examines trends in the European Union (EU) market for statutory audits of public interest entities (PIEs) from 2019 to 2021, with a focus on market concentration, audit quality risks, and the effectiveness of audit committees.
-This report is the third check-in on how the European Union's (EU’s) auditing standards and market structures are evolving since the *Article 27 of Regulation (EU) No 537/2014* [-@EU_Regulation_537_2014] was introduced.
+This report is the third check-in on how the EU’s auditing standards and market structures are evolving since the *Article 27 of Regulation (EU) No 537/2014* [-@EU_Regulation_537_2014] was introduced, which introduces a host of specific audit requirements for PIEs.
 A statutory audit is a legally mandated evaluation of financial records aimed at certifying the accuracy of companies’ financial statements, thereby enhancing stakeholder confidence and supporting the orderly functioning of markets [@EC_Auditing_Website].
@@
 
-Figure 1 illustrates the market shares of audit firms — Big 4, CR4, and 10KAP — for the number of PIE statutory audits conducted in 2021.
+@fig-market-shares illustrates the market shares of audit firms — Big 4, CR4, and 10KAP — for the number of PIE statutory audits conducted in 2021.
 Countries are sorted in a specific order reflecting the replication of Figure 3 from the European Commission’s 2024 report.
@@
 
-# Display the figure
-fig
+# Render the figure explicitly
+plt.figure(fig.number)
+plt.show()
 ```
@@
 # Research Design Choices and Assumptions {#sec-research_design_assumptions}
-The aim of Assignment III is to replicate a specific empirical table (Table 2 Panel A) from the seminal paper by @Leuz_2003.
-This table involves calculating the EM measures for firms across various countries over a defined period and examining the relationship between these measures and investor protection.
-The replication process includes data loading, preparation, cleaning, and normalization, followed by the application of statistical methods to compute and interpret financial metrics.
-For Assignment III, I pulled data from the Worldscope database through WRDS and used the Python programming language to carry out the empirical analysis.
+The aim of Assignment I is to replicate a specific empirical figure (Figure 3) from the report by @EC_Report_2024.
+This figure involves analyzing the market shares of audit firms (Big 4, CR4, and 10KAP) across EU countries and at the EU level for PIE statutory audits conducted in 2021.
+The replication process includes data loading, preparation, and cleaning, followed by the computation and visualization of market concentration measures.
+For Assignment I, I pulled data from the Audit Analytics database via WRDS and used the Python programming language to carry out the empirical analysis.
 Visual Studio Code was used as the Integrated Development Environment (IDE) for writing, debugging, and optimizing the Python code.
 
-The replication is based on data pulled from the Worldscope database, specifically from the `wrds_ws_company` and `wrds_ws_funda` tables, which were merged for the analysis.
-The first table provides company profile information, including items such as ISIN, Worldscope Identifiers, company name, and the country where the company is domiciled [@WRDS_WS_Company_2024].
-The latter table contains fundamental annual data at the company-year level, including items such as total assets, net income, and other relevant financial variables [@WRDS_WS_Funda_2024].
+The replication is based on data pulled from the Audit Analytics Transparency Reports database, specifically from `audit_europe.feed76_transparency_reports` table, which was filtered and aggregated for the analysis.
+The table offers Transparency Reports published annually by audit firms based in the EEA and Switzerland, along with the names of all entities listed in each audit firm’s Transparency Reports [@WRDS_Audit_Analytics_Transparency_Reports].
 
-Following @Leuz_2003, I focus the analysis on companies across various countries, ensuring that the data accurately reflects the fiscal years 1990 to 1999 as specified in the original study.
+Following @EC_Report_2024, I focus the analysis on statutory audits conducted for the fiscal year 2021, ensuring that the data accurately reflects the audit market structure in the EU as specified in the original report.
 The replication aims to mirror the research design as closely as possible with the available data.
 
-In addition, I impose the following assumptions to ensure clarity and consistency where the paper by @Leuz_2003 does not provide explicit guidance:
+In addition, I impose the following assumptions to ensure clarity and consistency where the report by @EC_Report_2024 does not provide explicit guidance:
 
-1.	The original study references the November 2000 version of the Worldscope Database.
-However, the data used for this analysis represents the latest available version, updated in July 2024, with quarterly frequency updates [@WRDS_2024].
-Due to potential adjustments and updates made to the database since November 2000, there may be differences between the databases that could affect the results.
-For example, companies may restate financials after the original reporting period, so these restatements are reflected in the later database version rather than the historical one.
-Moreover, the data vendor Refinitiv regularly updates its databases to correct errors and add new information, which may be included in the later data but not in the November 2000 snapshot.
+1.	The original report references the reports prepared by national competent authorities (NCAs) responsible for audit oversight as the main source of data, where the data refers to the years 2020 and 2021 and was collected in 2022 [@EC_Report_2024, p.1].
+However, this replication uses data obtained from the Audit Analytics Transparency Reports database via WRDS, which provides the latest available version as of September 2024, updated quarterly [@WRDS_Audit_Analytics_Transparency_Reports].
+Due to potential adjustments and updates made to the database since the original data collection period, there may be differences that could affect the results.
+For example, audit firms may disclose updated information in subsequent reports, which would be reflected in the later database version rather than the historical snapshot used in the original report.
+Furthermore, the data vendor regularly updates its databases to correct errors and add new information, which may be included in the later data but not in the 2022 snapshot.
 2.	The original paper outlines key terms that will be used in this project to ensure consistency and accuracy in the replication.
-@Leuz_2003 define earnings management as the manipulation of a firm’s reported economic performance by insiders to deceive certain stakeholders or to affect contractual outcomes.
-The authors describe investor protection as a key institutional factor that limits insiders’ acquisition of private control benefits, thereby reducing their incentives to manage accounting earnings by ensuring strong and well-enforced rights for outside investors.
-Finally, private control benefits are the advantages that insiders can gain from controlling a firm, which can include financial gains or other benefits not shared with other stakeholders [@Leuz_2003].
-3.	While pulling the data for analysis, I encountered negative values for some key financial metrics such as operating income (`item1250`) and net income before preferred dividends (`item1651`).
-The paper by @Leuz_2003 does not explicitly specify how to handle negative values in key financial metrics.
-For the purpose of this replication study, I included negative values in the analysis.
-This ensures that the analysis captures the full spectrum of earnings management activities across different countries.
-4.	Another potential source of discrepancies between the original and replicated tables may be the choice of variables pulled from Worldscope.
-For example, in the `wrds_ws_funda table`, both `item1151` and `item4051` are labeled as “DEPRECIATION, DEPLETION AND AMORTIZATION” [@WRDS_WS_Funda_2024].
-I chose `item1151` from the Income Statement rather than `item4051` from the Cash Flow Statement, based on the *Worldscope Balancing Model - Industrials* by Wharton Research Data Services [-@WRDS_Balancing_Model_2024].
-Since the authors do not specify the choice of variables used from the database, this could cause differences in the results.
-5.	The EM measures are based on scaled variables (e.g., operating cash flow scaled by lagged total assets).
-As such, the currency of the relevant data items should not affect the results as long as the same currency is used consistently for both the numerator and the denominator.
-This approach ensures comparability across different countries, regardless of their local currencies.
-Additionally, according to a document by @Thomson_Financial_2007 [p.20], all Worldscope data is consistently reported in the local currency of each firm’s country of domicile, eliminating the need for currency conversions in this project.
+@EC_Report_2024 defines the 'Big Four' as Deloitte, EY, KPMG, and PwC; the 'CR4' as the four largest audit firms in each country; and the '10KAP' as a group consisting of Baker Tilly, BDO, Deloitte, EY, Grant Thornton, KPMG, Mazars, Nexia, RSM, and PwC.
+3.	As for the geographical scope, I use the variable `trans_report_auditor_state` to identify the country associated with each audit firm rather than the audited company’s headquarter country (`headquarter_country` variable).
+Since the objective is to analyze audit market concentration and the market share of audit firms within specific countries, the country of the audit firm’s reported activity is more relevant than the headquarter country of the entities they audit.
+4.	Another potential source of discrepancies between the original and replicated tables may be the choice of variables pulled from Audit Analytics.
+For example, in the `audit_europe.feed76_transparency_reports` table, I chose `entity_map_fkey` over `entity_map_key`, as this unique identifier is created by WRDS for merging [@WRDS_Audit_Analytics_Transparency_Reports].
+Since the report does not specify the choice of variables used from the database, this could cause differences in the results.
+5.	For the purpose of replicating Figure 3, I use the `report_year` variable as the primary filter to define the relevant data for year 2021.
+This choice is based on the fact that `report_year` represents the idealized fiscal year of the Transparency Report, based on the *Europe Transparency Reports Data Dictionary* by Wharton Research Data Services [-@Europe_Transparency_Reports_Data_Dictionary], which aligns directly with the year-based aggregation needed for Figure 3.
+While the `report_period_end_date` provides the exact end date of each reporting period [@Europe_Transparency_Reports_Data_Dictionary], using `report_year` allows for a consistent, annual aggregation without requiring additional alignment of various period end dates.
+This approach simplifies the data extraction and ensures that all entries correspond to the fiscal year 2021.
 
-By following the steps provided in @sec-replication_steps and adhering to the assumptions made, I successfully replicated the analysis and produced the required table.
+By following the steps provided in @sec-replication_steps and adhering to the assumptions made, I successfully replicated the analysis and produced the required figure.
 A thorough step-by-step approach, with each step clearly documented, helped to understand and verify the outputs.
@@
 
-## Step 1: Pulling the Data and Managing the Databases
-Unlike Assignment I, where the data was provided externally, Assignment III involves pulling data directly from the Worldscope database, merging relevant tables, and preparing the data for further analysis from raw data to final output.
+## Step 1: Pulling the Data and Managing the Database
+This Assignment involves pulling data directly from the Audit Analytics database and preparing the data for further analysis from raw data to final output.
 
-To ensure data integrity, I filtered out rows with empty `item6105` (Worldscope Permanent ID) values, as this identifier is crucial for firm/year-level filtering in the data preparation step.
-A total of 125 observations from the dynamic data and 10,306 observations from the static data were removed.
-The dynamic and static datasets were then merged on the `item6105` identifier, representing the unique Worldscope Permanent ID.
-According to @WRDS_Overview_2024, this identifier should be used consistently within Worldscope data, as it remains stable over time.
-An inner join was used for this merge to ensure that only complete and consistent data from both tables were retained.
-
-Additionally, a filter was applied to retain only company rows.
-This was achieved by selecting rows where the `item6100` field equals ‘C’, indicating that the Worldscope Identifier represents a company.
-This step ensures that the analysis includes only company data, excluding averages, exchange rates, securities, or stock indices, as advised by @WRDS_Overview_2024.
-
-In line with the methodology used by @Leuz_2003, financial institutions were removed from the analysis based on their Standard Industrial Classification (SIC) codes.
-This was done by filtering out rows where the `item7021` identifier, representing the SIC code, falls within the range of 6000 to 6999.
-This ensures that the dataset focuses solely on non-financial companies.
-
-Further filtering ensured that only data from the 31 countries specified in the paper were included. These steps were applied to reduce the dataset size and improve the workflow.
-
-Moreover, the configuration file utilized additional refinements as proposed by @WRDS_Overview_2024.
-Specifically, using "A" (Annual) in the `freq` variable ensured that the data represents financial information reported on an annual basis, consistent with the methodology outlined in the paper by @Leuz_2003.
-This excludes data reported on a current, daily, or quarterly basis.
-After retaining only relevant company data, filtering out financial institutions, and focusing on the specified countries, the processed data was saved to a CSV file in the `data/pulled` directory, as specified in configuration file `config/pull_data_cfg`​.
+To ensure data relevance, the pulling process was restricted to data for the year 2021.
+The variable `trans_report_auditor_state` was used to filter the data to include only audit firms associated with the specific EU and EEA countries, as specified in configuration file `config/pull_data_cfg​`.
+This filtering step ensured that the analysis focuses on audit market concentration within the geographical scope and year range as defined by @EC_Report_2024.
 
 ## Step 2: Data Preparation {#sec-step2_preparation}
-To verify the pulled data, I checked the dataset for duplicate firm-year observations based on the combination of the Worldscope Permanent ID (`item6105`) and the fiscal year (`year_`).
-I confirmed that no duplicates were present, ensuring the accuracy of the data for further analysis.
-As the original study does not specify the net income measure, I used `item1651` (Net Income before Preferred Dividends) as the variable for net income.
-This choice aligns with the final net income figure reported in the income statement, based on *Worldscope Balancing Model - Industrials* by Wharton Research Data Services [-@WRDS_Balancing_Model_2024].
+The raw Transparency Report dataset initially contained 13,385 observations, representing entries for the year 2021.
+To ensure consistency with @EC_Report_2024, the abbreviation 'GR' for Greece was replaced with 'EL' in the trans_report_auditor_state field, affecting 188 rows.
 
-To prepare the data sample in line with the methodology outlined in the paper, it was essential to follow the requirements set by @Leuz_2003.
-First, countries with sufficient firm-year observations were filtered.
-Each country needed at least 300 firm-year observations for key accounting variables, including total assets, sales, net income, and operating income.
-In this step, no countries were eliminated, as all countries met the requirement, which aligns with the paper’s overview of countries (31 countries in total).
+The `number_of_disclosed_pies` variable provides the clearest indication of statutory audits conducted by each auditor for PIEs.
+The definition of PIEs varies across EU and EEA countries [@AccountancyEurope_PIE_2017], but if Transparency Reports consistently follow country-specific definitions, using the provided variable is reasonable.
+An important verification step involved comparing the reported `number_of_disclosed_pies` with the count of unique `entity_map_fkey` entries for each `transparency_report_fkey`.
+This analysis revealed that 5,696 transparency reports included fewer entities in the dataset than the disclosed PIEs, suggesting that some PIE entities might be missing from the dataset.
+This discrepancy reflects missing PIE data rather than the inclusion of non-PIE entities, hence it is assumed that all entities in the dataset are PIEs, maintaining consistency in the framework for analysis.
 
-In the second filtration step, firms with adequate consecutive data were identified.
-Each firm had to have income statement and balance sheet information for at least three consecutive years, with all key accounting variables mentioned in previous step being present.
-If a firm had at least three consecutive years of complete data at any point, all its data entries were retained in the final dataset.
-Therefore, only those countries and firms that met these criteria for all specified variables were retained in the dataset.
+Additionally, missing values in key fields, including `transparency_report_fkey`, `entity_map_fkey`, `auditor_fkey`, and `trans_report_auditor_state`, were checked, and it was confirmed that no rows needed to be removed due to missing data, ensuring the completeness of the dataset.
+To focus on relevant data, rows with `number_of_disclosed_pies` equal to or less than zero were filtered, but no such rows were identified, and all observations were retained.
 
-During the preparation step, 8,265 firms and 20,521 firm-year observations (all due to the second filtration step) were dropped, resulting in a final dataset with 18,040 non-financial firms and 123,469 firm-year observations.
-The differences in the numbers between the prepared dataset and the figures mentioned in the paper (70,955 firm-year observations and 8,616 non-financial firms) could be due to assumptions outlined in @sec-research_design_assumptions, such as variations in initial datasets, data updates, and filtering criteria.
-However, the original paper might have included additional data cleaning steps not explicitly mentioned, such as handling outliers, specific industry exclusions, or other criteria, which could affect the final counts.
+Duplicate entries in the `entity_map_fkey` column were then analyzed, revealing 1,011 duplicate entities.
+I assume that duplicate entries in the dataset do not introduce significant issues for the replication of Figure 3.
+Duplicate entries for the same entity are expected in cases where an entity is audited by multiple audit firms or networks, potentially as part of a joint audit.
+Since these duplicates retain consistent information on `trans_report_auditor_state`, they do not distort country-level distributions.
+Additionally, while duplicates might expand representation across different auditor networks or groups, the essence of Figure 3 — focusing on aggregate distributions — remains unaffected.
+Therefore, these duplicates were not removed.
 
-Given that the number of observations in this project is significantly higher than in the original study, Table 1 from @Leuz_2003 was partially replicated (focusing only on columns related to countries and firm-year observations) to highlight specific discrepancies that could arise for certain countries.
-So, the results in Table 2 show a higher number of firm-year observations for almost all countries compared to the paper by @Leuz_2003, likely due to database updates leading to a more comprehensive dataset.
-The trend in the number of firm-year observations across countries is consistent with the original study, with larger capital markets like Japan, the United Kingdom, and the United States having the highest numbers.
-The distribution of observations in my dataset shows higher mean and median values compared to the paper by @Leuz_2003, indicating a greater overall number of firm-year observations, while the minimum values are relatively close, suggesting that the overall distribution of observations across countries remains broadly similar.
+Auditor network names were standardized to ensure consistency, with names like `|Mazars Worldwide|Praxity Global Alliance|` simplified to `|Mazars Worldwide|`, for example.
+This standardization helped avoid double-counting and unify same entries under consistent naming conventions.
 
-However, there are significant discrepancies, particularly for the United States, which has 41,585 observations compared to 3,792 in the original study.
-@Leuz_2003 focused only on S&P 500 companies for the U.S., while this analysis uses a more comprehensive dataset from the Worldscope database, which includes detailed financial statement data and profile data on public companies globally.
-This broader scope results in a higher number of firm-year observations for the United States.
-Despite the differences in absolute numbers, the relative trends across countries align with the original paper.
+Audit firms were then categorized into groups, including the Big 4 (Deloitte, EY, KPMG, and PwC), the 10KAP (which includes the Big 4 along with other major networks listed in @sec-research_design_assumptions), 'Unaffiliated' firms (those outside the Big 4 and 10KAP).
+The `auditor_network` field contained 636 observations with missing values, which were grouped under the “Other (Blank)” category.
+I assume that these blank entries do not belong to the Big Four or 10KAP networks but may still contribute to the CR4 category in some countries.
+So, these entries were retained in the dataset and reviewed during the aggregation stage to validate their relevance.
 
-## Step 3: Analysis Implementation and Table Reproduction
-In this replication step, the calculation of the EM measures is implemented according to the methodology described by @Leuz_2003.
-The study employs four distinct proxies that capture different dimensions of earnings management, such as earnings smoothing and accrual manipulations.
-These measures are used to assess the pervasiveness of earnings management across countries.
-This step ensures that all statistical methods and groupings are accurately replicated, following the paper’s approach to capture the extent to which corporate insiders use accounting discretion to obscure their firm’s true economic performance.
+These preparation steps ensured alignment with the methodology of @EC_Report_2024 and the dataset’s readiness for subsequent analysis.
 
-### Smoothing Reported Operating Earnings Using Accruals (EM1) {#sec-EM1}
-EM1 is a measure that assesses how much insiders (e.g., company executives) smooth or stabilize reported earnings by manipulating accruals.
-The concept is that firms may attempt to reduce fluctuations in their reported operating income to present a more stable financial performance to investors.
-This measure reflects the degree to which insiders use accounting discretion to smooth reported earnings, with lower values of EM1 suggesting more extensive earnings smoothing.
+## Step 3: Analysis Implementation and Figure Reproduction
+The analysis step computes market concentration metrics for audit firms across EU countries and at the EU level using the prepared transparency data.
+The analysis begins by calculating the market shares of the Big 4, 10KAP, and CR4 audit firm groups for the number of statutory audits conducted in 2021.
+For each country, the Big 4 market share is derived as the percentage of PIE statutory audits conducted by Deloitte, EY, KPMG, and PwC, while the CR4 market share represents the top four audit firms in each country based on audit counts.
+The 10KAP market share includes the Big 4 along with six additional networks, providing a broader view of market concentration.
 
-To calculate this measure:
-
-1.	Compute Accruals by using changes in various components of a firm’s balance sheet and income statement, such as total current assets, cash, current liabilities, short-term debt, taxes payable, and depreciation:
-
+The following formula is used to calculate market share:
 $$
-\begin{aligned}
-\text{Accruals}_{it} &= (\Delta \text{CA}_{it} - \Delta \text{Cash}_{it}) - (\Delta \text{CL}_{it} - \Delta \text{STD}_{it} - \Delta \text{TP}_{it}) - \text{Dep}_{it} 
-& \\
-\text{where:} \\
-\Delta \text{CA}_{it} & = \text{Change in total current assets} \\
-\Delta \text{Cash}_{it} & = \text{Change in cash and cash equivalents} \\
-\Delta \text{CL}_{it} & = \text{Change in total current liabilities} \\
-\Delta \text{STD}_{it} & = \text{Change in short-term debt included in current liabilities} \\
-\Delta \text{TP}_{it} & = \text{Change in income taxes payable} \\
-\text{Dep}_{it} & = \text{Depreciation and amortization expense for firm i in year t} 
-\end{aligned}
+\text{Market Share (\%)} = \left( \frac{\text{Number of PIE Audits by Network Group}}{\text{Total PIE Audits in Country}} \right) \times 100
 $$
 
-2.	Derive Operating Cash Flow (CFO) that is indirectly derived from operating income minus accruals:
-$$
-\text{CFO}_{it} = \text{Operating Income}_{it} - \text{Accruals}_{it}
-$$
-
-3.	Calculate EM1, which is determined as the country’s median ratio of the firm-level standard deviations of operating income and operating cash flow.
-Both standard deviations are scaled by lagged total assets to control for differences in firm size:
-$$
-\text{EM1} = \text{Median}\left(\frac{\frac{\sigma(\text{Operating Income}_{it})}{\text{Total Assets}_{i(t-1)}}}{\frac{\sigma(\text{Operating Cash Flow}_{it})}{\text{Total Assets}_{i(t-1)}}}\right)
-$$
-
-### Smoothing and the Correlation Between Changes in Accounting Accruals and Operating Cash Flows (EM2)
-EM2 is designed to capture the extent to which insiders use accounting discretion to smooth reported earnings by examining the correlation between changes in accruals and changes in operating cash flows.
-A stronger negative correlation between these variables indicates that accruals are being used to buffer the impact of cash flow fluctuations, potentially masking the true economic performance of the firm.
-
-To calculate this measure:
-
-1. Compute Accruals and CFO using same Formulas used in @sec-EM1.
-
-2. Calculate the change in Accruals and CFO for each firm-year:
-$$
-\Delta \text{Accruals}_{it} = \text{Accruals}_{it} - \text{Accruals}_{i(t-1)}
-$$
-$$
-\Delta \text{CFO}_{it} = \text{CFO}_{it} - \text{CFO}_{i(t-1)}
-$$
-
-3. Scale both changes by lagged total assets:
-$$
-\frac{\Delta \text{Accruals}_{it}}{\text{Total Assets}_{i(t-1)}}, \quad \frac{\Delta \text{CFO}_{it}}{\text{Total Assets}_{i(t-1)}}
-$$
-
-4.
-Calculate EM2, which is computed as the Spearman correlation between these scaled changes, aggregated at the country level over the pooled set of firms within each country, meaning that the correlation reflects the aggregated relationship between changes in accruals and changes in CFO across all firms in that country:
-$$
-\text{EM2} = \rho\left(\frac{\Delta \text{Accruals}_{it}}{\text{Total Assets}_{i(t-1)}}, \frac{\Delta \text{CFO}_{it}}{\text{Total Assets}_{i(t-1)}}\right)
-$$
-
-### Discretion in Reported Earnings: The Magnitude of Accruals (EM3)
-EM3 is a measure that obtains the degree of discretion that insiders exercise when reporting earnings, particularly when it comes to misstating a firm’s economic performance.
-It focuses on the magnitude of accruals as a proxy for earnings management.
-Insiders can use accruals to overstate or understate earnings, thereby achieving certain financial targets or presenting a skewed view of the firm’s performance.
-Higher scores of EM3 indicate more earnings management, as larger magnitudes of accruals relative to cash flow from operations suggest greater discretion in financial reporting.
-
-To calculate this measure:
-
-1. Compute Accruals and CFO using same Formulas used in @sec-EM1.
-
-2. Calculate the absolute values of accruals and CFO for each firm-year.
-
-3.
-EM3 is calculated as the country-level median ratio of the absolute value of firms’ accruals to the absolute value of their cash flow from operations.
-This ratio is used as it scales the magnitude of accruals to a measure of the firm’s size and performance, allowing for comparison across firms of different sizes and performance levels:
-$$
-\text{EM3} = \text{Median}\left(\frac{|\text{Accruals}_{it}|}{|\text{CFO}_{it}|}\right)
-$$
-
-### Discretion in Reported Earnings: Small Loss Avoidance (EM4)
-EM4 captures the extent to which firms manage their reported earnings to avoid small losses.
-This is an important measure that quantifies the propensity of firms to avoid small losses through earnings management.
-By calculating the ratio of small profits to small losses, it provides a clear indication of how widespread this practice is within a particular country.
-A higher EM4 value suggests that firms in a given country are more likely to use accounting discretion to convert small losses into small profits, reflecting a tendency toward earnings management.
-Note that I used `item1651` (Net Income before Preferred Dividends) as the variable for net earnings, as mentioned in @sec-step2_preparation.
-I also ensured that each country included in the EM4 calculation had at least five small loss observations, as per the methodology outlined by @Leuz_2003.
-The review found that no countries were excluded from the analysis due to this criterion.
-
-EM4 is calculated as the ratio of the number of small profits to the number of small losses for each country, where small profits and small losses are determined by scaling net earnings by lagged total assets:
-$$
-\text{EM4} = \frac{\text{Small Profits}}{\text{Small Losses}}
-$$
-$$
-\begin{alignedat}{2}
-\text{where:} \\ \quad & \text{Small Profits} && = \frac{\text{Net Earnings}_{it}}{\text{Total Assets}_{it-1}} \in [0, 0.01] \\
-& \text{Small Losses} && = \frac{\text{Net Earnings}_{it}}{\text{Total Assets}_{it-1}} \in [-0.01, 0)
-\end{alignedat}
-$$
-
-### Aggregate Measure of Earnings Management
-The aggregate earnings management score is a summary measure that captures the overall level of earnings management in each country.
-This measure is calculated by ranking countries on each of the four individual earnings management measures (EM1-EM4), where a higher rank indicates a greater degree of earnings management.
-The final aggregate score is the average of these ranks, providing a comprehensive view of a country’s earnings management practices.
-The countries are sorted in descending order based on their aggregate score.
-
-To calculate this measure:
-$$
-\text{Aggregate Earnings Management Score}_{\text{country}} = \frac{\text{Rank}_{\text{EM1}} + \text{Rank}_{\text{EM2}} + \text{Rank}_{\text{EM3}} + \text{Rank}_{\text{EM4}}}{4}
-$$
+The analysis also includes the creation of a bar chart visualizing market shares for all countries and the EU aggregate, saved in both PNG and pickle formats.
+The aggregated market share data is saved in CSV format, ensuring reproducibility and further analysis opportunities.
 
 # Results {#sec-results}
-The comparison between the calculated EM1 values and those reported in the original paper reveals a strong alignment, with only minor differences observed across most countries.
-The overall pattern of EM1 values is consistent, effectively capturing the underlying trends in earnings management across different countries.
-Any differences, such as those for the Philippines, Taiwan, and Norway, are relatively small and likely due to variations in data sources, updates, or slight methodological differences, as outlined in @sec-research_design_assumptions.
-The summary statistics (mean, median, min, max, and standard deviation) also closely match those in the original study, further supporting the robustness of the replication process.
-The results indicate that Anglo-American countries tend to have higher EM1 values, suggesting less earnings smoothing compared to Continental Europe and Asia, consistent with the conclusion of @Leuz_2003 that earnings are smoother in Continental Europe and Asia than in Anglo-American countries after controlling for the volatility of cash flows.
+To verify the results of analysis step and observe trends before plotting, I have additionally saved the CSV output in `aggregated_market_shares.csv`.
+The analysis of market shares for statutory audits across European countries reveals distinct patterns of market concentration and auditor dominance.
+Big 4 firms exhibit a significant presence, often commanding a dominant share in many countries.
+For example, countries such as the Czech Republic, Estonia, and Finland demonstrate Big 4 market shares exceeding 95%, reflecting the significant influence of these firms in specific markets.
+However, the extent of this dominance varies, with lower Big 4 shares observed in countries like Bulgaria (17.5%) and Greece (43.6%), indicating a more diversified audit market in these regions.
+The analysis reveals that the Big 4 hold a market share exceeding 80% in 11 EU countries.
+This aligns with rthe finding of @EC_Report_2024, which similarly highlights that the Big 4 dominate over 80% of the market share in 11 Member States.
 
-The replication findings regarding EM2 indicate that earnings smoothing is more pronounced in countries like Greece and Japan, while countries like Canada and the United States exhibit less pervasive smoothing behavior, in line with the findings of @Leuz_2003.
-Despite some differences in magnitude, the overall trends are consistent, reinforcing the idea that earnings smoothing varies significantly across different countries.
-In fact, these small variations do not significantly alter the overall conclusions.
-Specifically, the minimum and maximum EM2 values in the replication closely match the original findings, with Greece showing the most earnings smoothing (min: -0.942 vs.
--0.928) and Norway the least (max: -0.786 vs.
--0.722).
-This consistency reinforces the accuracy of the replication.
+The inclusion of additional audit networks through the 10KAP grouping increases the market share substantially across all countries, as expected.
+This is particularly evident in countries like Ireland, Malta, and the Netherlands, where the 10KAP share reaches 100%, reflecting the full inclusion of statutory audits under this grouping.
+In contrast, countries like Bulgaria and Poland, with 10KAP shares of 41.8% and 63.6%, respectively, suggest the presence of a substantial number of statutory audits conducted by unaffiliated or smaller firms.
 
-The replicated EM3 results are very similar to those of the original study, with close alignment in summary statistics and consistent identification of South Africa and Germany as the countries with the minimum and maximum EM3 values.
-Additionally, the replication confirms that the magnitude of firms' accruals relative to their operating cash flows is smaller in the U.K.
-and the U.S.
-compared to Austria, Germany, and South Korea, consistent with the original findings.
+In the analysis of CR4 market shares, I verified whether the four largest audit firms (CR4) in each country overlapped entirely with the Big 4 audit firms.
+This overlap occurred in 15 out of 28 countries, where CR4 and Big 4 market shares were identical.
+In the remaining 13 countries, at least one non-Big 4 firm contributed significantly to the statutory audits, resulting in distinct CR4 market shares.
+These findings align with @EC_Report_2024, which notes that in 13 Member States, the Big Four are not the four largest audit firms in terms of the total number of PIE statutory audit opinions.
+This highlights the gradual diversification in the audit market in some countries, where non-Big 4 firms are playing a more prominent role, while the Big 4 continue to dominate in others.
 
-The replicated results show significant regional differences in EM4 values, reflecting varying degrees of loss avoidance.
-Spain has the highest EM4 value, highlighting strong loss avoidance among European firms.
-Norway, with the lowest EM4 value, corresponds to the minimum value in the original paper, confirming less aggressive loss avoidance.
-High EM4 values in Asian countries like Indonesia (5.667) and India (6.000) further support the trend that European and Asian firms exhibit greater loss avoidance than their Anglo-American counterparts, such as the United States (1.912) and Canada (2.402), which have lower EM4 values.
-These results align with the original study's conclusions on regional loss avoidance behavior.
+At the EU level, the aggregated market shares further underscore the dominance of these major players.
+The Big 4 account for 70% of statutory audits, compared to @EC_Report_2024, which notes an average EU market share of 59% for the Big 4 in 2021, down from 70% in 2018.
+The broader 10KAP grouping in my output reaches nearly 90%, exceeding the 81% reported by @EC_Report_2024.
+The CR4 market share at the EU level also stands at 70%, aligning with the Big 4, and is consistent with the average CR4 market share reported by @EC_Report_2024.
+These results highlight the varying degrees of market concentration across Europe, with certain countries demonstrating a highly centralized audit market dominated by the Big 4, while others display a more distributed market landscape, incorporating both major and smaller audit firms.
+The slight discrepancies in the figures suggest potential differences in methodologies, data coverage, or sample definitions.
 
-The comparison between the original and replicated aggregate earnings management scores reveals a strong overall alignment with consistent key trends.
-In the original study, Austria and Greece are at the top of the rankings with aggregate scores of 28.3, indicating the highest levels of earnings management.
-In the replication, Greece and Austria remain at the top with slightly adjusted scores of 27.5 and 27.2, respectively.
-Taiwan and South Korea also rank high, mirroring their positions in the original study.
-While some differences in specific rankings and scores exist, such as the United States consistently having the lowest level of earnings management (2.0 in the original and 3.5 in the replication), this slight variation could be attributed to differences in the datasets, particularly due to the use of a more comprehensive dataset from the Worldscope database, which includes financial data on U.S.
-public companies globally, in contrast to only S&P index firms used in the original paper, as discussed in @sec-step2_preparation.
-In general, these variations are minor and do not undermine the overall consistency.
+Based on the results for Austria, the Big 4 market share is 86.7% in my analysis, aligning closely with the approximate 85% represented in bar chart of @EC_Report_2024.
+Similarly, the 10KAP market share of 92.55% in my analysis corresponds well with the slightly higher bar in the report, and the CR4 market share of 86.7% aligns perfectly with the Big 4 share, confirming the dominance of these firms in Austria’s audit market.
+This consistency highlights the robustness of the replicated analysis for this country.
 
-In summary, the replication effectively captures the original study's key findings, confirming that European and Asian countries generally exhibit higher levels of earnings management than their Anglo-American counterparts.
-The strong similarity between the original and replicated results underscores the robustness and reliability of the findings.
+The results for Romania show notable discrepancies.
+The EC Report indicates market shares of 15% for the Big 4, 16% for CR4, and 24% for 10KAP, while my analysis reports significantly higher values: 63%, 70.61%, and 78.38%, respectively.
+These differences likely arise from assumptions made in @sec-research_design_assumptions like variations in PIE definitions and auditor classifications.
+The report by @EC_Report_2024 may use stricter criteria, reflecting Romania’s decentralized market where non-Big 4 firms dominate, unlike the broader inclusion in my dataset.
+
+In summary, the replication effectively captures the original report's key findings, confirming the dominance of the Big 4 and the broader 10KAP group in the European audit market.
+The analysis highlights varying degrees of market concentration across countries, with some demonstrating a more decentralized audit landscape.
+While discrepancies in individual country results, such as Romania, underline the importance of methodological alignment, the strong similarity between the overall market shares reported in the original report and the replicated results underscores the robustness and reliability of the findings.
 
 # Conclusion {#sec-conclusion}
-The project effectively demonstrates the use of an open science and collaborative workflow for analyzing earnings management measures across various countries.
-By following a systematic approach and utilizing the TRR 266 Template for Reproducible Empirical Accounting Research, I successfully replicated an empirical table from @Leuz_2003 and provided insights into trends in earnings management practices, which, similar to the original study, vary significantly across countries.
+The project effectively demonstrates the use of a systematic and collaborative workflow for analyzing audit market concentration across Europe, leveraging the TRR 266 Template for Reproducible Empirical Accounting Research.
+By following a structured approach, I successfully replicated a key figure and empirical findings of @EC_Report_2024, offering insights into the dominance of the Big 4 and the broader 10KAP group in the European audit market.
 
-@Leuz_2003 document systematic differences in the level of earnings management across 31 countries, suggesting that insiders use earnings management to conceal firm performance and protect their private control benefits.
-The study further argues that stronger investor protection reduces earnings management by limiting insiders’ ability to secure private benefits.
-By conducting a descriptive cluster analysis following the descriptive statistics replicated in this project, the paper identifies groupings of countries with similar institutional characteristics, demonstrating that earnings management varies systematically across these institutional clusters.
+This assignment is an exact replication of the report by @EC_Report_2024, where the goal was to closely follow the report’s approach wherever possible and make justified assumptions when necessary.
+While exact replications often yield different samples and outcomes due to dataset updates or methodological variations, my results align closely with the statistics presented in the original study, reinforcing the reliability of the replication process.
 
-The replicated results reveal significant differences in earnings management practices across countries, with consistent patterns in the relative magnitudes of the individual measures.
-These differences are particularly notable in areas such as loss avoidance and earnings smoothing.
-In line with the original study, the replication suggests that firms in countries with more developed equity markets, dispersed ownership structures, strong investor rights, and robust legal enforcement tend to engage in less earnings management.
-This consistent finding across both the original and replicated studies underscores the critical role of the institutional environment in shaping corporate earnings management behaviors on an international scale.
+The analysis reveals significant variation in market concentration across countries, with the Big 4 accounting for over 80% of market share in 11 countries, while others show a more decentralized landscape where non-Big 4 firms contribute substantially.
+These findings align closely with the original study results, confirming the reliability of the data and methodologies used in this replication.
+However, discrepancies in specific countries, such as Romania, highlight the importance of transparent methodologies and consistent definitions when conducting such analyses.
 
-The replication analysis also reveals that the individual earnings management measures are highly correlated, resulting in similar rankings across the four measures and the aggregate earnings management score.
-This consistency supports the original study’s conclusion that a single underlying factor likely represents the four individual measures.
-As in the original results, the replicated country rankings based on the aggregate score show high ranks for countries like Greece, Austria, and South Korea, and lower ranks for countries such as Australia, the U.K., and the U.S.
-This reinforces the robustness of the aggregate measure in capturing the overall level of earnings management across different countries.
-
-This final assignment required the comprehensive application of all the skills learned and feedback received from previous assignments, making it a more complex project workflow.
+This assignment required a comprehensive application of skills learned throughout the course, incorporating data analysis, replication techniques, and visualization.
 In the future, this repository can be cloned or forked (if made public) to kickstart further projects on earnings management measures analysis.
-Additionally, the included presentation template can be rendered alongside the paper for further academic use, providing a visually engaging summary of the research findings.
 Thanks for reading!

### 2024-11-18 code/python/do_analysis old.py

@@ -1,266 +1,0 @@
-# We start by loading the libraries that we will use in this analysis.
-import pickle
-import pandas as pd
-from utils import read_config, setup_logging
-
-# Set up logging
-log = setup_logging()
-
-def main():
-    log.info("Performing main analysis...")
-    
-    # Load the configuration file
-    cfg = read_config('config/do_analysis_cfg.yaml')
-    
-    # Load the prepared financial data using the path from the config file
-    financial_data = load_data(cfg['prepared_data_save_path'])
-
-    # Calculate EM1
-    country_em1, summary_stats_em1 = calculate_em1(financial_data)
-
-    # Calculate EM2
-    country_em2, summary_stats_em2 = calculate_em2(financial_data)
-
-    # Calculate EM3
-    country_em3, summary_stats_em3 = calculate_em3(financial_data)
-    
-    # Calculate EM4
-    country_em4, summary_stats_em4 = calculate_em4(financial_data)
-
-    # Calculate Aggregate Earnings Management Score and create final table
-    final_table = calculate_aggregate(country_em1, country_em2, country_em3, country_em4)
-
-    # Create the final combined table with both metrics and summary statistics
-    final_combined_table = create_final_combined_table(final_table)
-
-    # Save the final combined table using the path from the config file
-    save_results(final_combined_table, cfg['results'])
-
-    log.info("Performing main analysis...Done!")
-
-def load_data(data_path):
-    """
-    Load the prepared financial data from the specified path.
-    """
-    df = pd.read_csv(data_path)
-    return df
-
-def calculate_em1(df):
-    """
-    Calculate EM1 for each firm and then take the country-level median.
-    This function combines all necessary steps into one.
-    """
-    # Step 1: Calculate Accruals
-    # Group by unique firm (item6105) and calculate diffs within each firm
-    # If a firm does not report information on taxes payable or short-term debt,
-    # then the change in both variables is assumed to be zero (per paper).
-    df['delta_CA'] = df.groupby('item6105')['item2201'].diff()  # Change in total current assets
-    df['delta_Cash'] = df.groupby('item6105')['item2003'].diff()  # Change in cash and cash equivalents
[…]

### 2024-11-18 code/python/do_analysis.py

@@ -30,7 +30,7 @@ def main():
     save_market_shares(market_shares, cfg['aggregated_data_save_path'])
 
     # Plot the results
-    plot_market_shares(market_shares, cfg['figure_save_path'])
+    plot_market_shares(market_shares, cfg['figure_save_path'], cfg['pickle_save_path'])
     
     log.info("Performing main analysis and plotting...Done!")
 
@@ -205,7 +205,7 @@ def save_market_shares(df, save_path):
     df.to_csv(save_path, index=False)
     log.info(f"Market shares saved to {save_path}.")
 
-def plot_market_shares(market_shares, save_path):
+def plot_market_shares(market_shares, save_path, pickle_path):
     """
     Create a patterned bar chart for Big 4, CR4, and 10KAP market shares by country (including EU),
     and save the figure as both PNG and pickle formats.
@@ -246,7 +246,6 @@ def plot_market_shares(market_shares, save_path):
     # Add labels, title, and legend
     ax.set_xlabel('Country', fontsize=12)
     ax.set_ylabel('Market Share (%)', fontsize=12)
-    ax.set_title("Replication of Figure 3 from EC Report (2024): Audit Firms' Market Share in Number of PIE Statutory Audits (2021)", fontsize=14)
     ax.set_xticks([i + bar_width for i in x])
     ax.set_xticklabels(countries, rotation=90)
     
@@ -260,7 +259,7 @@ def plot_market_shares(market_shares, save_path):
     log.info(f"Figure saved to {save_path}.")
     
     # Save the figure as a pickle file
-    pickle_path = save_path.replace('.png', '.pickle')
+    os.makedirs(os.path.dirname(pickle_path), exist_ok=True)
     with open(pickle_path, 'wb') as f:
         pickle.dump(fig, f)
     log.info(f"Figure saved as pickle to {pickle_path}.")

### 2024-11-18 code/python/prepare_data old.py

@@ -1,110 +1,0 @@
-# --- Header -------------------------------------------------------------------
-# Prepare the pulled data for further analysis as per the requirements by Leuz et al.
-#
-# (C) Melisa Mazaeva -  See LICENSE file for details
-# ------------------------------------------------------------------------------
-
-import pandas as pd
-import pickle
-from utils import read_config, setup_logging
-
-log = setup_logging()
-def main():
-    log.info("Preparing data for analysis ...")
-    cfg = read_config('config/prepare_data_cfg.yaml')
-
-    # Load the pulled data
-    wrds_data = pd.read_csv(cfg['worldscope_sample_save_path'])
-    initial_obs_count_pulled = len(wrds_data)
-    initial_firm_count_pulled = len(wrds_data['item6105'].unique())
-    log.info(f"Initial number of observations after pulling data: {initial_obs_count_pulled}")
-    log.info(f"Initial number of firms after pulling data: {initial_firm_count_pulled}")
-
-    # Check for duplicate firm-year observations
-    dup_obs = wrds_data[wrds_data.duplicated(subset=['item6105', 'year_'], keep=False)]
-    if not dup_obs.empty:
-        log.warning(f"Found {dup_obs.shape[0]} duplicate firm-year observations. Removing duplicates.")
-        wrds_data = wrds_data.drop_duplicates(subset=['item6105', 'year_'], keep='first')
-
-    # Filter countries with at least 300 firm-year observations for key accounting variables as in paper
-    filtered_countries_data, eliminated_countries = filter_countries(wrds_data)
-
-    # Print eliminated countries
-    if eliminated_countries:
-        log.info(f"Countries eliminated after filtration: {', '.join(eliminated_countries)}")
-    else:
-        log.info("No countries were eliminated after filtration.")
-
-    # Filter firms with at least three consecutive years of income statement and balance sheet information as in paper
-    initial_firm_count = len(filtered_countries_data['item6105'].unique())
-    initial_obs_count = len(filtered_countries_data)
-    
-    filtered_firms_data = filter_firms(filtered_countries_data)
-
-    final_firm_count = len(filtered_firms_data['item6105'].unique())
-    final_obs_count = len(filtered_firms_data)
-
-    firms_dropped = initial_firm_count - final_firm_count
-    obs_dropped = initial_obs_count - final_obs_count
-
[…]

### 2024-11-18 code/python/pull_wrds_data old.py

@@ -1,98 +1,0 @@
-# --- Header -------------------------------------------------------------------
-# See LICENSE file for details
-#
-# This code pulls data from WRDS Audit Analytics Database 
-# ------------------------------------------------------------------------------
-import os
-from getpass import getpass
-import dotenv
-
-import pandas as pd
-from utils import read_config, setup_logging
-import wrds
-
-log = setup_logging()
-
-def main():
-    '''
-    Main function to pull data from WRDS.
-
-    This function reads the configuration file, gets the WRDS login credentials, and pulls the data from WRDS.
-
-    The data is then saved to a csv file.
-    '''
-    cfg = read_config('config/pull_data_cfg.yaml')
-    wrds_login = get_wrds_login()
-    wrds_data = pull_wrds_data(cfg, wrds_login)
-    wrds_data.to_csv(cfg['worldscope_sample_save_path'], index=False)
-
-
-def get_wrds_login():
-    '''
-    Gets the WRDS login credentials.
-    '''
-    if os.path.exists('secrets.env'):
-        dotenv.load_dotenv('secrets.env')
-        wrds_username = os.getenv('WRDS_USERNAME')
-        wrds_password = os.getenv('WRDS_PASSWORD')
-        return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
-    else:
-        wrds_username = input('Please provide a wrds username: ')
-        wrds_password = getpass(
-            'Please provide a wrds password (it will not show as you type): ')
-        return {'wrds_username': wrds_username, 'wrds_password': wrds_password}
-
-def pull_wrds_data(cfg, wrds_authentication):
-    '''
-    Pulls WRDS access data.
-    '''
-    db = wrds.Connection(
-        wrds_username=wrds_authentication['wrds_username'], wrds_password=wrds_authentication['wrds_password']
-    )
-
-    log.info('Logged on to WRDS ...')
-
-    dyn_var_str = ', '.join(cfg['dyn_vars'])
-    stat_var_str = ', '.join(cfg['stat_vars'])
-
-    log.info("Pulling dynamic Worldscope data ... ")
-    wrds_data_dynamic = db.raw_sql(
-        f"SELECT {dyn_var_str} FROM tr_worldscope.wrds_ws_funda WHERE {cfg['cs_filter']}"
-    )
-    log.info("Pulling dynamic Worldscope data ... Done!")
-
-    log.info("Pulling static Worldscope data ... ")
[…]
